---
title: "EHV-1 Gene dynamics"
author: Balázs Kakuk
output: 
  html_document:
    theme: cosmo
    toc: yes
    toc_float:
      collapsed: false
---

```{css, echo=FALSE}
    body .main-container {
      max-width: 90% !important;
      width: 90% !important;
    }
    body {
      max-width: 90% !important;
      margin-left: auto;
      margin-right: auto;
    }
```


```{r include=FALSE}
## hrbrthemes::ipsum:

knitr::opts_chunk$set(fig.retina=2, echo = FALSE, fig.align = 'center', message = F, warning = F)
```

```{r ipsum_setup, message=FALSE, warning=FALSE, cache=FALSE, echo=FALSE}
library(prettydoc)
library(hrbrthemes, quietly = T)
#library(GenomicFeatures, quietly = T)
library(DESeq2, quietly = T)
library(ggpubr, quietly = T)
#library(misc, quietly = T)
library(dplyr, quietly = T)
#library(tidyverse, quietly = T)
library(ggsci, quietly = T)
#library(Gviz, quietly = T)
library(dplyr, quietly = T)
library(tidyr)
library(scales)
#library(plyr)
library(moanin, quietly = T)
library(grid, quietly = T)
library(gridExtra, quietly = T) 
library(ggpubr, quietly = T)
library(data.table)
library(ggh4x)
library(knitr)
library(formattable)


###
source('_WF.part0.R')

#### Settings ####

orf.mean.average_genome  <- NULL
orf.mean.viral_read.count  <- NULL
orf.mean.total_read.count  <- NULL
orf.mean.host_read.count  <- NULL
orf.mean.raw_read.count  <- NULL
orf.sum.average_genome  <- NULL
orf.sum.viral_read.count  <- NULL
orf.sum.total_read.count  <- NULL
orf.sum.host_read.count  <- NULL
orf.sum.raw_read.count  <- NULL

orf.perc.mean.TSS.only  <- NULL
orf.perc.mean.TES.only  <- NULL
orf.perc.mean.TSS.TES  <- NULL


orf.perc.sum.TSS.only  <- NULL
orf.perc.sum.TES.only  <- NULL
orf.perc.sum.TSS.TES  <- NULL


## Miscallenaous
palette <-  colorvec[-c(9,13)] #pal_npg()(10)

writetables <- T


```


# Combine all Gene counting methods

*Now we combine the TSS-only, the TES-only and the TSS.TES methods for all normalization methods*


## Project info
*Samples: dcDNA Seq*  
*pychopper: no*  
*mapping version: v6*  
*LoRTIA: yes (stranded only)*  



# TSS only

## Only those reads were counted, where LoRTIA   
found the *5-prime* adapters!

```{r message=FALSE, warning=FALSE, cache=FALSE, echo=FALSE}
## Main
#save.data <- 'PRV.rebase.RData'
EndType <- 'TSS'

project_config  <- fread('project_config.txt')
outdir  <- project_config$outdir
res.dir <- outdir; try({ dir.create(res.dir) })
fig.dir <- paste0(res.dir, '/', EndType, '_abund'); # try({ dir.create(fig.dir) })## save plots to this directory



```

```{r, message=FALSE, warning=FALSE, cache=FALSE, echo=FALSE}
#### Import results of all Normalization methods

##
try({

##
norm.method   <- 'viral_read.count'
orf.perc.mean <- fread(paste0(fig.dir, '.norm_LoRTIA', '/', norm.method, '_', EndType, '_abund.mean.tsv'))
orf.perc.mean[,norm.method := norm.method]
orf.mean.viral_read.count <- orf.perc.mean

orf.sum.perc  <- fread(paste0(fig.dir, '.norm_LoRTIA', '/', norm.method, '_', EndType, '_abund.tsv'))
orf.sum.perc[,norm.method := norm.method]
orf.sum.viral_read.count <- orf.sum.perc
})

##
try({
norm.method   <- 'host_read.count'
orf.perc.mean <- fread(paste0(fig.dir, '.norm_host', '/', norm.method, '_', EndType, '_abund.mean.tsv'))
orf.perc.mean[,norm.method := norm.method]
orf.mean.host_read.count <- orf.perc.mean

orf.sum.perc <- fread(paste0(fig.dir, '.norm_host', '/', norm.method, '_', EndType, '_counts.norm.tsv'))
orf.sum.perc[,norm.method := norm.method]
orf.sum.host_read.count <- orf.sum.perc
})

##
try({
norm.method   <- 'VST'
orf.perc.mean <- fread(paste0(fig.dir, '.norm_VST', '/', norm.method, '_', EndType, '_abund.mean.tsv'))
orf.perc.mean[,norm.method := norm.method]
orf.mean.host_read.count <- orf.perc.mean

orf.sum.perc <- fread(paste0(fig.dir, '.norm_VST', '/', norm.method, '_', EndType, '_counts.norm.tsv'))
orf.sum.perc[,norm.method := norm.method]
orf.sum.host_read.count <- orf.sum.perc
})

###
orf.perc.mean <- rbind(
  #orf.mean.average_genome,
  orf.mean.viral_read.count
  #orf.mean.total_read.count,
 # orf.mean.host_read.count,
  #orf.mean.raw_read.count
)

orf.perc.mean[,Kinetic_class := factor(Kinetic_class, levels = c('IE', 'IE/E', 'E', 'IE/E/L', 'E/L', 'L', 'unknown'))]

###
orf.perc.sum <- rbind(
  #orf.sum.average_genome,
  orf.sum.viral_read.count
  #orf.sum.total_read.count,
  #orf.sum.host_read.count,
  #orf.sum.raw_read.count
)

orf.perc.sum[,Kinetic_class := factor(Kinetic_class, levels = c('IE', 'IE/E', 'E', 'IE/E/L', 'E/L', 'L', 'unknown'))]

orf.perc.mean[, EndType := EndType]
orf.perc.sum[,  EndType := EndType]

orf.perc.mean.TSS.only <- orf.perc.mean
orf.perc.sum.TSS.only  <- orf.perc.sum

```


# TES only

## Only those reads were counted, where LoRTIA   
found the *3-prime* adapters!

```{r message=FALSE, warning=FALSE, cache=FALSE, echo=FALSE}
## Main
#save.data <- 'PRV.rebase.RData'
EndType <- 'TES'

project_config  <- fread('project_config.txt')
outdir  <- project_config$outdir
res.dir <- outdir; try({ dir.create(res.dir) })
fig.dir <- paste0(res.dir, '/', EndType, '_abund'); # try({ dir.create(fig.dir) })## save plots to this directory



```

```{r, message=FALSE, warning=FALSE, cache=FALSE, echo=FALSE}
#### Import results of all Normalization methods

##
try({

##
norm.method   <- 'viral_read.count'
orf.perc.mean <- fread(paste0(fig.dir, '.norm_LoRTIA', '/', norm.method, '_', EndType, '_abund.mean.tsv'))
orf.perc.mean[,norm.method := norm.method]
orf.mean.viral_read.count <- orf.perc.mean

orf.sum.perc  <- fread(paste0(fig.dir, '.norm_LoRTIA', '/', norm.method, '_', EndType, '_abund.tsv'))
orf.sum.perc[,norm.method := norm.method]
orf.sum.viral_read.count <- orf.sum.perc
})

##
try({
norm.method   <- 'host_read.count'
orf.perc.mean <- fread(paste0(fig.dir, '.norm_host', '/', norm.method, '_', EndType, '_abund.mean.tsv'))
orf.perc.mean[,norm.method := norm.method]
orf.mean.host_read.count <- orf.perc.mean

orf.sum.perc <- fread(paste0(fig.dir, '.norm_host', '/', norm.method, '_', EndType, '_counts.norm.tsv'))
orf.sum.perc[,norm.method := norm.method]
orf.sum.host_read.count <- orf.sum.perc
})

##
try({
norm.method   <- 'VST'
orf.perc.mean <- fread(paste0(fig.dir, '.norm_VST', '/', norm.method, '_', EndType, '_abund.mean.tsv'))
orf.perc.mean[,norm.method := norm.method]
orf.mean.host_read.count <- orf.perc.mean

orf.sum.perc <- fread(paste0(fig.dir, '.norm_VST', '/', norm.method, '_', EndType, '_counts.norm.tsv'))
orf.sum.perc[,norm.method := norm.method]
orf.sum.host_read.count <- orf.sum.perc
})

###
orf.perc.mean <- rbind(
  #orf.mean.average_genome,
  orf.mean.viral_read.count
  #orf.mean.total_read.count,
 # orf.mean.host_read.count,
  #orf.mean.raw_read.count
)

orf.perc.mean[,Kinetic_class := factor(Kinetic_class, levels = c('IE', 'IE/E', 'E', 'IE/E/L', 'E/L', 'L', 'unknown'))]

###
orf.perc.sum <- rbind(
  #orf.sum.average_genome,
  orf.sum.viral_read.count
  #orf.sum.total_read.count,
  #orf.sum.host_read.count,
  #orf.sum.raw_read.count
)

orf.perc.sum[,Kinetic_class := factor(Kinetic_class, levels = c('IE', 'IE/E', 'E', 'IE/E/L', 'E/L', 'L', 'unknown'))]


orf.perc.mean[, EndType := EndType]
orf.perc.sum[,  EndType := EndType] 

orf.perc.mean.TES.only <- orf.perc.mean
orf.perc.sum.TES.only  <- orf.perc.sum

```




# Canonic TSS.TES

## Only those reads were counted, where LoRTIA   
found *both 5-prime AND 3-prime* adapters!

```{r message=FALSE, warning=FALSE, cache=FALSE, echo=FALSE}
## Main
#save.data <- 'PRV.rebase.RData'
EndType <- 'TSS.TES'

project_config  <- fread('project_config.txt')
outdir  <- project_config$outdir
res.dir <- outdir; try({ dir.create(res.dir) })
fig.dir <- paste0(res.dir, '/', EndType, '_abund'); # try({ dir.create(fig.dir) })## save plots to this directory



```

```{r, message=FALSE, warning=FALSE, cache=FALSE, echo=FALSE}
#### Import results of all Normalization methods

##
try({

##
norm.method   <- 'viral_read.count'
orf.perc.mean <- fread(paste0(fig.dir, '.norm_LoRTIA', '/', norm.method, '_', EndType, '_abund.mean.tsv'))
orf.perc.mean[,norm.method := norm.method]
orf.mean.viral_read.count <- orf.perc.mean

orf.sum.perc  <- fread(paste0(fig.dir, '.norm_LoRTIA', '/', norm.method, '_', EndType, '_abund.tsv'))
orf.sum.perc[,norm.method := norm.method]
orf.sum.viral_read.count <- orf.sum.perc
})

##
try({
norm.method   <- 'host_read.count'
orf.perc.mean <- fread(paste0(fig.dir, '.norm_host', '/', norm.method, '_', EndType, '_abund.mean.tsv'))
orf.perc.mean[,norm.method := norm.method]
orf.mean.host_read.count <- orf.perc.mean

orf.sum.perc <- fread(paste0(fig.dir, '.norm_host', '/', norm.method, '_', EndType, '_counts.norm.tsv'))
orf.sum.perc[,norm.method := norm.method]
orf.sum.host_read.count <- orf.sum.perc
})

##
try({
norm.method   <- 'VST'
orf.perc.mean <- fread(paste0(fig.dir, '.norm_VST', '/', norm.method, '_', EndType, '_abund.mean.tsv'))
orf.perc.mean[,norm.method := norm.method]
orf.mean.host_read.count <- orf.perc.mean

orf.sum.perc <- fread(paste0(fig.dir, '.norm_VST', '/', norm.method, '_', EndType, '_counts.norm.tsv'))
orf.sum.perc[,norm.method := norm.method]
orf.sum.host_read.count <- orf.sum.perc
})

###
orf.perc.mean <- rbind(
  #orf.mean.average_genome,
  orf.mean.viral_read.count
  #orf.mean.total_read.count,
 # orf.mean.host_read.count,
  #orf.mean.raw_read.count
)

orf.perc.mean[,Kinetic_class := factor(Kinetic_class, levels = c('IE', 'IE/E', 'E', 'IE/E/L', 'E/L', 'L', 'unknown'))]

###
orf.perc.sum <- rbind(
  #orf.sum.average_genome,
  orf.sum.viral_read.count
  #orf.sum.total_read.count,
  #orf.sum.host_read.count,
  #orf.sum.raw_read.count
)

orf.perc.sum[,Kinetic_class := factor(Kinetic_class, levels = c('IE', 'IE/E', 'E', 'IE/E/L', 'E/L', 'L', 'unknown'))]

orf.perc.mean[, EndType := EndType]
orf.perc.sum[,  EndType := EndType]

orf.perc.mean.TSS.TES  <- orf.perc.mean
orf.perc.sum.TSS.TES   <- orf.perc.sum

```



# Combine dynamics
```{r}
orf.perc.mean <- rbind(
  orf.perc.mean.TSS.only,
  orf.perc.mean.TES.only,
  orf.perc.mean.TSS.TES)


orf.perc.sum <- rbind(
  orf.perc.sum.TSS.only,
  orf.perc.sum.TES.only,
  orf.perc.sum.TSS.TES)

```


```{r, message=FALSE, warning=FALSE, cache=FALSE, echo=FALSE}
### Summarise based on literature kinetic class
kin.class.sum  <- orf.perc.sum[,.(sum_read_count=sum(read_count), mean_read_count=mean(read_count), sd_read_count=sd(read_count)), 
                              by=.(EndType, Kinetic_class, hpi, Time, cell_line, seqnames, norm_base, sample, group, norm.method)]


### Summarise based on de-novo kinetic clusters
cluster.sum <- orf.perc.sum[,.(sum_read_count=sum(read_count), mean_read_count=mean(read_count), sd_read_count=sd(read_count)), 
                              by=.(EndType, cluster, hpi, Time, cell_line, seqnames, norm_base, sample, group, norm.method)]




###
norm.method <- 'All_Methods_Combined'
```


```{r message=FALSE, warning=FALSE, cache=FALSE, echo=FALSE}
## Main
#save.data <- 'PRV.rebase.RData'
EndType <- 'All_Methods_Combined'

project_config  <- fread('project_config.txt')
outdir  <- project_config$outdir
res.dir <- outdir; try({ dir.create(res.dir) })
fig.dir <- paste0(res.dir, '/', EndType, '_abund'); try({ dir.create(fig.dir) })## save plots to this directory


```


## Gene dynamics, according to literature-based kinetic classes {.tabset}

```{r, fig.show='hold',  fig.width = 30, fig.height = 4, eval=T}
plotfun <- function(DT, 
                    geom1 = geom_pointrange(aes(ymin = mean-sd, ymax = mean+sd, colour = cell_line)),
                    geom2 = geom_line(aes(x=Time, y=mean, color=cell_line)),
                    #geom2 = geom_smooth(aes(x=Time, y=mean, color=cell_line)),
                    ncol = 10, title = NULL, #"Normalized ORF counts", 
                    strip.backgr = pal_d3()(10)[2],
                    ...) {
  
  if (nrow(DT) == 0) { 
    ggp <- NULL } else { # try({ 

    ggp <- 
      ggplot(DT[, ],
             aes(Time, mean, color = cell_line)
                     ) + 
      geom1 +
      geom2 + 
      #scale_color_manual(values = palette) +
      #scale_y_continuous(labels = scales::percent_format()) +
      scale_x_continuous(breaks = unique(DT$Time)) +
      theme_ipsum() +
      theme(plot.margin = unit(c(10,1,1,1), 'mm'),
            legend.position = 'none',
            ...) +
      facet_nested(cols=vars(EndType, norm.method, gene),
                   scales = "free_y") #, rows=vars(gene))
      
      if(!is.null(title)) { ggp <- ggp + ggtitle(title) }
    
  
  }   # })
  
  return(ggp)

}


figw     <- 24
figh     <- 18

length(unique(orf.perc.mean[, gene])) == length(unique(orf.perc.sum[, gene]))
## OK!

DT <- orf.perc.mean
normMethods <- c('viral_read.count','host_read.count')

genestoplot <- unique(orf.perc.mean[, gene])
genetoplot <- 'ORF1'
pattern <- paste0("\\b", genetoplot, "\\b")

grep(pattern, genestoplot, value = T)


mapfun <- function(i, genestoplot, normMethods, DT, palette, ...) {
  genetoplot <- genestoplot[i]
  pattern <- paste0("\\b", genetoplot, "\\b")
  DTsub <- DT[norm.method %in% normMethods & grepl(pattern, gene)]
  gg <- plotfun(DTsub, title = genetoplot, ...) + scale_color_manual(values = palette[i])
  ggsave(paste0(fig.dir, '/', genetoplot, '.jpg'), 
         gg, height = 9, width = 24)
  gg
}

colorvec <- rep(na.omit(colorvec), 2)

mapfun(4, unique(orf.perc.mean$gene), c('viral_read.count','host_read.count'), orf.perc.mean, colorvec)

ggsave('TEST.jpg', height = 24, width = 24)

```


### Mean and SD with linear connection
```{r, fig.show='hold',  fig.width = 30, fig.height = 360, eval=T}
plotlist <- purrr::map(1:length(unique(orf.perc.mean$gene)), mapfun,
                       unique(orf.perc.mean$gene), c('viral_read.count','host_read.count'), orf.perc.mean, colorvec)

cowplot::plot_grid(plotlist = plotlist, ncol=1)

```

### Mean and SD with loess function
```{r, fig.show='hold',  fig.width = 30, fig.height = 240, eval=F}
plotlist <- purrr::map(1:length(unique(orf.perc.mean$gene)), mapfun,
                       unique(orf.perc.mean$gene), c('viral_read.count','host_read.count'), orf.perc.mean, colorvec,
                       geom2 = geom_smooth(aes(x=Time, y=mean, color=cell_line)))

cowplot::plot_grid(plotlist = plotlist, ncol=1)

```


### Barplots
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}
figw     <- 24
figh     <- 18

```


```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=F}
top       <- 10

orf.perc.mean[, rank := frank(-mean, ties.method = "average"), by = .(hpi, Time, cell_line, norm.method)]

top_genes <- unique(orf.perc.mean[,.(hpi, Time, cell_line, norm.method, rank, gene, Kinetic_class)])

#dcast(top_genes[cell_line == 'C6' & norm.method == 'WO' & Time == 4,], 
#      hpi + Time + cell_line + norm.method ~ rank, value.var = 'gene')

# there are many ties

top_genes <- unique(top_genes[rank <= top, .(gene, rank)])
top_genes <- unique(top_genes$gene)

orf.perc.top <- orf.perc.mean[!gene %in% top_genes, .(mean = sum(mean), sd=mean(sd)), by=.(hpi, Time, cell_line, norm.method, Kinetic_class)]
orf.perc.top$gene <- NA
orf.perc.top$Kinetic_class <- NA

orf.perc.top <- rbind(orf.perc.top, 
                      orf.perc.mean[gene %in% top_genes,.(hpi, Time, cell_line, norm.method, gene, Kinetic_class, mean, sd)]
                      )

plotbar <- function(DT, title, subtitle) {

  ggp <- ggplot(DT) +
    geom_col(aes(x = cell_line, y = mean, fill = gene), color='black') +
    ggtitle( title, subtitle) + 
    coord_flip() +
    scale_fill_manual(values = na.omit(colorvec)) +
    theme_bw() +
    theme(#strip.text.y = element_blank(),
          strip.backgr = pal_d3()(10)[1]) +
    facet_nested(rows=vars(Time, cell_line), cols=vars(Kinetic_class), scales = "free_y")
  
  ggp

}

title <- paste0('Top ', top, ' genes')

ggp <- suppressWarnings(plotbar(orf.perc.top[norm.method == 'average_genome'], subtitle = 'Average viral genome coverage', title=title) )
ggsave(paste0(fig.dir, '/', EndType, '_average_genome_Bar.jpg'), ggp, width = figw, height = figh, limitsize = F)

ggp <- suppressWarnings(plotbar(orf.perc.top[norm.method == 'viral_read.count'], subtitle = 'Viral read count', title=title) )
ggsave(paste0(fig.dir, '/', EndType, '_viral_read.count_Bar.jpg'), ggp, width = figw, height = figh, limitsize = F)

ggp <- suppressWarnings(plotbar(orf.perc.top[norm.method == 'total_read.count'], subtitle = 'Total read count', title=title) )
ggsave(paste0(fig.dir, '/', EndType, '_total_read.count_Bar.jpg'), ggp, width = figw, height = figh, limitsize = F)

ggp <- suppressWarnings(plotbar(orf.perc.top[norm.method == 'host_read.count'], subtitle = 'Host read count', title=title) )
ggsave(paste0(fig.dir, '/', EndType, '_host_read.count_Bar.jpg'), ggp, width = figw, height = figh, limitsize = F)

ggp <- suppressWarnings(plotbar(orf.perc.top[norm.method == 'WO'], subtitle = 'Without normalization', title=title) )
ggsave(paste0(fig.dir, '/', EndType, '_WO_Bar.jpg'), ggp, width = figw, height = figh, limitsize = F)
```

### Dotplots
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}
figw     <- 18
figh     <- 12

```


```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=F}
plotdot <- function(DT,  title, subtitle) {
  ggp <- ggplot(DT) +
    geom_point(aes(x = Time, y = mean, fill = cell_line), color='black', shape = 21) +
    geom_line(aes(x = Time, y = mean, color = cell_line)) +
    ggtitle( title, subtitle) + 
    #coord_flip() +
    scale_fill_manual(values = na.omit(colorvec)) +
    theme_bw() +
    theme(#strip.text.y = element_blank(),
          strip.backgr = pal_d3()(10)[1]) +
    facet_nested_wrap(~gene, ncol=8, scales = "free_y")
  
  ggp

}

ggp <- suppressWarnings(plotdot(orf.perc.top[norm.method == 'average_genome'], subtitle = 'Average viral genome coverage', title=title) )
ggsave(paste0(fig.dir, '/', EndType, '_average_genome_Dot.jpg'), ggp, width = figw, height = figh, limitsize = F)

ggp <- suppressWarnings(plotdot(orf.perc.top[norm.method == 'viral_read.count'], subtitle = 'Viral read count', title=title) )
ggsave(paste0(fig.dir, '/', EndType, '_viral_read.count_Dot.jpg'), ggp, width = figw, height = figh, limitsize = F)

ggp <- suppressWarnings(plotdot(orf.perc.top[norm.method == 'total_read.count'], subtitle = 'Total read count', title=title) )
ggsave(paste0(fig.dir, '/', EndType, '_total_read.count_Dot.jpg'), ggp, width = figw, height = figh, limitsize = F)

ggp <- suppressWarnings(plotdot(orf.perc.top[norm.method == 'host_read.count'], subtitle = 'Host read count', title=title) )
ggsave(paste0(fig.dir, '/', EndType, '_host_read.count_Dot.jpg'), ggp, width = figw, height = figh, limitsize = F)

ggp <- suppressWarnings(plotdot(orf.perc.top[norm.method == 'WO'], subtitle = 'Without normalization', title=title) )
ggsave(paste0(fig.dir, '/', EndType, '_WO_Dot.jpg'), ggp, width = figw, height = figh, limitsize = F)


```



## Combined dynamics of kinetic classes/clusters

```{r}
figw <- 20
figh <- 12
ncol <- 10
DT <- orf.perc.mean
normMethods <- c('viral_read.count','host_read.count')

```


### De Novo - Significant
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}
ggp <- 
      ggplot(cluster.sum[norm.method %in% normMethods],
             aes(Time, mean_read_count, color = as.factor(cluster))
                     #, add = 'loess'
                     ) + 
      geom_point() +
      geom_smooth() + 
      scale_color_manual(values = palette[]) +
      #scale_y_continuous(labels = scales::percent_format()) +
      theme_ipsum() +
      theme(plot.margin = unit(c(10,1,1,1), 'mm')) +
      #theme(strip.background.y = element_rect(fill = alpha(strip.backgr, 0.4))) +
      # theme(strip.text.y = element_blank()) +
      #ggtitle(title) +
      facet_nested(rows=vars(EndType, norm.method), cols=vars(cluster), scales = "free_y") ## cols=vars(norm.method)
      # facet_wrap(~ labels, scales = 'free_y', ncol = ncol)


ggp

ggsave(paste0(fig.dir, '/', EndType, '_', norm.method, '_Combined.DeNovo.Clusters.', '_points.loess.jpg'), ggp, width = figw, height = figh, limitsize = F)
```

```{r, fig.show='hold',  fig.width = 24, fig.height = 16, eval=F}
gg <- ggplot(orf.perc.sum[norm.method %in% normMethods], #[gene %in% c('ORF18', 'ORF19')], 
             aes(x=Time, y=read_count)) +
    #geom_line(aes(group=rep_gene, color=as.factor(best_label))) +
    geom_smooth(aes(group=rep_gene, color=as.factor(labels))) +
    #ggtitle(cluster_toplot) + 
    scale_color_manual(values = palette) +
    theme_ipsum() +
    facet_nested(cols=vars(norm.method, EndType),
                 scales = "free_y",
                 rows=vars(as.factor(cluster))
                 ) 
  

#gg
```

```{r, fig.show='hold',  fig.width = 24, fig.height = 16, eval=T}
DT <- orf.perc.sum
DT[,rep      := gsub('.*h_', '', sample)]
DT[,rep_gene := paste0(gene, '::', rep)]

mapfun <- function(i, clusters_toplot, cluster_name, normMethods, DT) {
  
  cluster_toplot <- clusters_toplot[i]
  
  DTsub <- as.data.frame(DT[norm.method %in% normMethods])
  
  DTsub <- as.data.table(DTsub[DTsub[,cluster_name] == cluster_toplot &
                               !is.na(DTsub[,cluster_name]), ])
  
  gg <- ggplot(DTsub ,
             aes(x=Time, y=read_count)) +
    #geom_line(aes(group=rep_gene, color=as.factor(best_label))) +
    geom_smooth(aes(group=rep_gene, color=as.factor(cluster))) +
    ggtitle(cluster_toplot) + 
    scale_color_manual(values = palette[i]) +
    theme_ipsum() +
    theme(legend.position = 'none') +
    facet_nested(rows=vars(norm.method, EndType),
                 scales = "free_y"
                 #,rows=vars(as.factor(best_label))
                 ) 
  gg
  
}

plotlist <- purrr::map(1:length(unique(na.omit(DT$cluster))), mapfun, unique(na.omit(DT$cluster)), 'cluster', c("viral_read.count", "host_read.count"), DT)

cowplot::plot_grid(plotlist=plotlist, nrow=1, align = 'vh', axis = 'tlbr')
  

```




### Literature clusters
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}
ggp <- 
      ggplot(kin.class.sum[norm.method %in% normMethods],
             aes(Time, mean_read_count, color = Kinetic_class)
                     #, add = 'loess'
                     ) + 
      geom_point() +
      geom_smooth() + 
      scale_color_manual(values = palette[]) +
      #scale_y_continuous(labels = scales::percent_format()) +
      theme_ipsum() +
      theme(plot.margin = unit(c(10,1,1,1), 'mm')) +
      #theme(strip.background.y = element_rect(fill = alpha(strip.backgr, 0.4))) +
      # theme(strip.text.y = element_blank()) +
      #ggtitle(title) +
      facet_nested(rows=vars(EndType, norm.method), cols=vars(Kinetic_class), scales = "free_y") ## cols=vars(norm.method)
      #facet_wrap(~ Kinetic_class, scales = 'free_y', ncol = ncol)


ggp

ggsave(paste0(fig.dir, '/', EndType, '_', norm.method, '_Combined.Literature.Clusters.', '_points.loess.jpg'), ggp, width = figw, height = figh, limitsize = F)
```

```{r, fig.show='hold',  fig.width = 24, fig.height = 16, eval=F}
gg <- ggplot(orf.perc.sum[norm.method %in% normMethods], #[gene %in% c('ORF18', 'ORF19')], 
             aes(x=Time, y=read_count)) +
    #geom_line(aes(group=rep_gene, color=as.factor(best_label))) +
    geom_smooth(aes(group=rep_gene, color=Kinetic_class)) +
    #ggtitle(cluster_toplot) + 
    scale_color_manual(values = palette) +
    #theme_ipsum() +
    facet_wrap(~norm.method + EndType + Kinetic_class, ncol=5,
                 scales = "free_y") 
  

#gg
```

```{r, fig.show='hold',  fig.width = 24, fig.height = 16, eval=T}
DT <- orf.perc.sum
DT[,rep      := gsub('.*h_', '', sample)]
DT[,rep_gene := paste0(gene, '::', rep)]

mapfun <- function(i, clusters_toplot, cluster_name, normMethods, DT) {
  
  cluster_toplot <- clusters_toplot[i]
  
  DTsub <- as.data.frame(DT[norm.method %in% normMethods])
  DTsub <- as.data.table(DTsub[DTsub[,cluster_name] == cluster_toplot, ])
  
  gg <- ggplot(DTsub ,
             aes(x=Time, y=read_count)) +
    #geom_line(aes(group=rep_gene, color=as.factor(best_label))) +
    geom_smooth(aes(group=rep_gene, color=as.factor(Kinetic_class))) +
    ggtitle(cluster_toplot) + 
    scale_color_manual(values = palette[i]) +
    theme_ipsum() +
    theme(legend.position = 'none') +
    facet_nested(rows=vars(norm.method, EndType),
                 scales = "free_y"
                 #,rows=vars(as.factor(best_label))
                 ) 
  gg
  
}

plotlist <- purrr::map(1:length(unique(DT$Kinetic_class)), mapfun, unique(DT$Kinetic_class), 'Kinetic_class', c("viral_read.count", "host_read.count"), DT)

cowplot::plot_grid(plotlist=plotlist, nrow=1, align = 'vh', axis = 'tlbr')
  

```





## Compare literature and de novo kinetic classes

*NA shows the number of genes where the clustering did not yield an unambiguous result.*
  
```{r, fig.show='hold',  fig.width = 18, fig.height = 14, eval=T}
setDT(orf.perc.mean)
class.dt <- unique(orf.perc.mean[, .(EndType, norm.method, gene, Kinetic_class, cluster)])
class.dt <- class.dt[, .N, by=.(EndType, norm.method, Kinetic_class, cluster)]

#cluster_labels

ggplot(class.dt[norm.method %in% normMethods],
       aes(x=Kinetic_class, y=N, fill=as.factor(cluster))) +
  geom_col(position = 'dodge') +
  scale_fill_manual(values = palette) +
  #scale_y_continuous(labels = scales::percent_format()) +
  theme_ipsum() +
  facet_nested(cols=vars(cluster), rows=vars(EndType, norm.method))


ggplot(class.dt[norm.method %in% normMethods],
       aes(x=as.factor(cluster), y=N, fill=Kinetic_class)) +
  geom_col(position = 'stack') +
  scale_fill_manual(values = palette) +
  #scale_y_continuous(labels = scales::percent_format()) +
  theme_ipsum() +
  facet_nested(cols=vars(cluster), rows=vars(EndType, norm.method))


```





## Gene-pair ratios

These pairs were analyzed: 
ie180/us1, ie180/ep0, us1/ep0, noir1/noir2, noir1/ie180, noir2/ie180

```{r, fig.show='hold',  fig.width = 10, fig.height = 7, eval=F}




ab.ORF65    <- setnames(orf.perc.sum[gene == 'ORF65_1',    .(hpi, Time, cell_line, sample, group, norm.method, read_count)], "read_count", 'ORF65')

ab.ORF64  <- setnames(orf.perc.sum[gene == 'ORF64_1',  .(hpi, Time, cell_line, sample, group, norm.method, read_count)], "read_count", 'ORF64')

ab.ORF63    <- setnames(orf.perc.sum[gene == 'ORF63',    .(hpi, Time, cell_line, sample, group, norm.method, read_count)], "read_count", 'ORF63')

#ab.noir   <- setnames(orf.perc.sum[gene == 'NOIR_1',   .(hpi, Time, cell_line, sample, group, norm.method, read_count)], "read_count", 'NOIR_1')

#ab.noir2  <- setnames(orf.perc.sum[gene == 'NOIR-2', .(hpi, Time, cell_line, sample, group, norm.method, read_count)], "read_count", 'NOIR-2')


ab.merged <- merge(ab.ORF65,    ab.ORF64,  by=c(metacols, 'norm.method'))
ab.merged <- merge(ab.merged, ab.ORF63,    by=c(metacols, 'norm.method'))
#ab.merged <- merge(ab.merged, ab.noir,   by=c(metacols, 'norm.method'))
#ab.merged <- merge(ab.merged, ab.noir2,  by=c(metacols, 'norm.method'))

ab.merged[, ORF64_ORF65   :=  ORF64   /  ORF65]
ab.merged[, ORF64_ORF63   :=  ORF64   /  ORF63]
ab.merged[, ORF65_ORF63     :=  ORF65     /  ORF63]
#ab.merged[, NOIR1_NOIR2 :=  NOIR    / `NOIR-2`]
#ab.merged[, NOIR_ORF64  :=  NOIR    /  ORF64]
#ab.merged[, NOIR2_ORF64 := `NOIR-2` /  ORF64]

#ab.merged[, ORF65_ORF64   :=  ORF65     / ORF64     ]
#ab.merged[, ORF63_ORF64   :=  ORF63     / ORF64     ]
#ab.merged[, ORF63_ORF65     :=  ORF63     / ORF65       ]
#ab.merged[, NOIR2_NOIR1 := `NOIR-2` / NOIR      ]
#ab.merged[, ORF64_NOIR  :=  ORF64   / NOIR_1      ]
#ab.merged[, ORF64_NOIR2 :=  ORF64   / `NOIR-2`  ]

gpairs <- c(
'ORF64_ORF65' , 
'ORF64_ORF63'  , 
'ORF65_ORF63' 
#'NOIR1_NOIR2' ,
#'NOIR_ORF64'  ,
#'NOIR2_ORF64' ,

#,'ORF65_ORF64'   ,
#,'ORF63_ORF64'   ,
#,'ORF63_ORF65'     ,
#,'NOIR2_NOIR1' ,
#;'ORF64_NOIR'
#,'ORF64_NOIR2' 
)
gpairs <- gsub('_', '/', gpairs)

ab.merged.gt <- ab.merged[,-c(7:9)]

ab.merged.gt <- melt(ab.merged.gt, variable.name = 'gene_pair', value.name = 'ratio', id.vars = c(1:6))

ab.merged.gt$gene_pair <- gsub('_', '/', ab.merged.gt$gene_pair)
ab.merged.gt[,gene_pair := factor(gene_pair, levels = gpairs)]

ab.merged.gt[,mean := mean(ratio),   by=.(hpi, Time, cell_line, group, norm.method, gene_pair)]
ab.merged.gt[,sd   := sd  (ratio),   by=.(hpi, Time, cell_line, group, norm.method, gene_pair)]
ab.merged.gt[,sum  := sum (ratio),   by=.(hpi, Time, cell_line, group, norm.method, gene_pair)]



plot_data <- ab.merged.gt[norm.method == 'viral_read.count'
                 # norm.method == 'host_read.count' 
                 , ]

ggmean <- ggplot(plot_data) + 
  #geom_point(aes(Time, ratio, color=cell_line)) +
  geom_line(aes(Time, mean, color=cell_line, group=cell_line)) +
  geom_pointrange(aes(Time, mean, color=cell_line, group=cell_line, ymin=mean-sd, ymax=mean+sd)) +
  theme_bw() +
  #facet_nested(rows = vars(norm.method), cols=vars(Time))
  facet_wrap(~gene_pair, scales = 'free', ncol=3)


ggmean
ggsave(paste0(fig.dir, '/gene_ratios_mean.jpg'), width = 10, height = 7)


ggrep <- ggplot(plot_data[# norm.method == 'viral_read.count'
                 # norm.method == 'host_read.count' 
                 , ]) + 
  geom_point(aes(Time, ratio, color=cell_line)) +
  geom_smooth(aes(Time, ratio, color=cell_line, group=cell_line)) +
  #geom_pointrange(aes(Time, mean, color=cell_line, group=cell_line, ymin=mean-sd, ymax=mean+sd)) +
  theme_bw() +
  #facet_nested(rows = vars(norm.method), cols=vars(Time))
  facet_wrap(~gene_pair, scales = 'free')


ggrep
ggsave(paste0(fig.dir, '/gene_ratios_rep.jpg'), width = 10, height = 7)


ggsum <- ggplot(plot_data) + 
  geom_point(aes(Time, sum, color=cell_line)) +
  geom_line(aes(Time, sum, color=cell_line, group=cell_line)) +
  #geom_pointrange(aes(Time, mean, color=cell_line, group=cell_line, ymin=mean-sd, ymax=mean+sd)) +
  theme_bw() +
  #facet_nested(rows = vars(norm.method), cols=vars(Time))
  facet_wrap(~gene_pair, scales = 'free', ncol=3)


ggsum
ggsave(paste0(fig.dir, '/gene_ratios_sum.jpg'), width = 10, height = 7)

```




```{r}
## Export data
if (writetables) {
  
  #fwrite(orf.perc.sum,  paste0(fig.dir, '/', norm.method, '_', EndType, '_kin.class.sum.tsv'), sep = '\t')  
  #fwrite(orf.perc.sum,  paste0(fig.dir, '/', norm.method, '_', EndType, '_sig.cluster.sum.tsv'), sep = '\t') 
  #fwrite(orf.perc.sum,  paste0(fig.dir, '/', norm.method, '_', EndType, '_all.cluster.sum.tsv'), sep = '\t') 
  fwrite(orf.perc.sum,  paste0(fig.dir, '/', norm.method, '_', EndType, '_counts.norm.tsv'), sep = '\t')
  fwrite(orf.perc.mean, paste0(fig.dir, '/', norm.method, '_', EndType, '_abund.mean.tsv'), sep = '\t')
  
}
```


```{r bib, include=FALSE}
# KEEP THIS AT THE END OF THE DOCUMENT TO GENERATE A LOCAL bib FILE FOR PKGS USED
knitr::write_bib(sub("^package:", "", grep("package", search(), value=TRUE)), file='skeleton.bib')
```
