---
title: "EHV-1 Genes' transcript ratios"
author: Bal치zs Kakuk
output: 
  html_document:
    theme: cosmo
    toc: yes
    toc_float:
      collapsed: false
---

```{css, echo=FALSE}
    body .main-container {
      max-width: 90% !important;
      width: 90% !important;
    }
    body {
      max-width: 90% !important;
      margin-left: auto;
      margin-right: auto;
    }
```


```{r include=FALSE}
## hrbrthemes::ipsum:

knitr::opts_chunk$set(fig.retina=2, echo = FALSE, fig.align = 'center', message = F, warning = F)
```

```{r ipsum_setup, message=FALSE, warning=FALSE, cache=FALSE, echo=FALSE}
library(prettydoc)
library(hrbrthemes, quietly = T)
#library(GenomicFeatures, quietly = T)
library(DESeq2, quietly = T)
library(ggpubr, quietly = T)
#library(misc, quietly = T)
library(dplyr, quietly = T)
#library(tidyverse, quietly = T)
library(ggsci, quietly = T)
#library(Gviz, quietly = T)
library(dplyr, quietly = T)
library(tidyr)
library(scales)
#library(plyr)
library(grid, quietly = T)
library(gridExtra, quietly = T) 
library(ggpubr, quietly = T)
library(data.table)
library(ggh4x)
library(knitr)
library(formattable)
library(ggrepel)
library(utils)

#library(moanin, quietly = T)
#library(stageR); library(edgeR) ; library(Biobase)
#library(limma) ; library(DEXSeq)

dontrun <- T
source('customclust.R')
source('mclust.R')


###
viral.ref <- "NC_001491.2.mRNA_corrected.v3.gff3"
source('_WF.part0.R')

#### Settings ####

## Main
#save.data <- 'PRV.rebase.RData'
EndType <- 'TX.ratio'

project_config  <- fread('project_config.txt')
outdir  <- project_config$outdir
res.dir <- outdir; try({ dir.create(res.dir) })
fig.dir <- paste0(res.dir, '/', EndType); try({ dir.create(fig.dir) })## save plots to this directory

## Miscallenaous
colorvec <- na.omit(c(
  colorvec[c(1:5,15,13,10,24,7,26,32,33)],
  colorvec[-c(1:5,15,13,10,24,7,26,32,33)]))

colorvec <- colorvec[-c(9,13)]


colorvec <- colorvec[-c(10,13,17,18)]

colorvec <- colorvec[c(1:2,14,3:length(colorvec))]


palette  <- colorvec #pal_npg()(10)

writetables <- T


#### For Gene/Transcript plotting

## Gene Feature annotation


genome.plotdata <- make.genome.plot.data(feature.df, feature.col = 8, feature.name = 'ID')
genome.plotdata$strand <- factor(genome.plotdata$strand, levels = c('+', '-', '*'))
gene.plotdata <- genome.plotdata
## GENE NAMES !
gene.plotdata$gene_name <- gsub('_.*', '', gene.plotdata$gene)
#gene.plotdata$start > gene.plotdata$end
## Misc
alpha <- data.frame(cov_geom        = 0.6,
                    gene_geom       = 0.75,
                    unstranded_geom = 0.75)



```

# Project info
*Samples: dcDNA Seq*  
*pychopper: no*  
*mapping version: v6*  
*LoRTIA: yes (stranded only)*  

*Filtering samples: C6_2h_3 & PC-12_8h_3*  

## Only those reads were counted, where LoRTIA found *both* 5- and 3-prime adapters!

```{r, eval=F}

TR.gff.compare.uni.CAGE     <- fread(paste0(outdir, '/TR.gff.compare.uni.CAGE.tsv'), na.strings = '')
TR.data <- fread(paste0(outdir, '/TR.data.tsv'), na.strings = '')
TR.reads.gfffile <- paste0(outdir, '/TR.reads.gff2')
TR.gff <- data.table(as.data.frame(rtracklayer::import.gff2(TR.reads.gfffile)))


```


```{r}
#### Import transcript and gene (cluster) counts
TR.gene.cluster <- fread(paste0(outdir, '/', 'TR.gene.cluster.tsv'), na.strings = '')
```

```{r, fig.width=16, fig.height=9}

TR.gene.cluster[,.N,class_code]

#TR.gene.cluster[,.N,distance_prime5]

ggplot(TR.gene.cluster, aes(distance_prime5)) + 
  geom_histogram(bins=100) + 
  geom_vline(xintercept = c(-10,10), color='red')  +
  geom_vline(xintercept = c(-2,2), color='blue') +
  facet_wrap(~class_code, scales = 'free')


```

### Count only reads with distance <= 2 as equal ('=')
```{r, fig.width=16, fig.height=9, eval=F}

TR.gene.cluster <- TR.gene.cluster[, 
    class_code := fifelse((distance_prime5 > 2 | distance_prime5 < -2 ) & class_code == '=', '~', class_code) ]

TR.gene.cluster[,.N,class_code]

ggplot(TR.gene.cluster, aes(distance_prime5)) + 
  geom_histogram(bins=100) + 
  geom_vline(xintercept = c(-10,10), color='red')  +
  geom_vline(xintercept = c(-2,2), color='blue') +
  facet_wrap(~class_code, scales = 'free')

```


```{r}
#### Import Toram et al annotation transferred from Kaplan
source('import.ref.TRs.R')

#### Reference annotation with CAGE results
TR.merged.data <- fread(paste0(outdir, '/TR.Ref.data.CAGE.tsv'))

#### Read-transcripts counts with adapter info
TR.adapt.count  <- fread(paste0(outdir, '/TR.adapt.count.tsv'), na.strings = '')

#### consider adapters?
# possible values : adapters <- 'both' | 'either' | 'any' | 'prime5' | 'prime3'
adapters <- 'both'
       if(adapters == 'both') {
  TR.adapt.filt <- TR.adapt.count[correct_tss == T & correct_tes == T]
  
} else if(adapters == 'either') {
  TR.adapt.filt <- TR.adapt.count[correct_tss == T | correct_tes == T, .(count = sum(count)), by=.(seqnames, strand, TR_ID, sample)]
  
} else if(adapters == 'any') {
  TR.adapt.filt <- TR.adapt.count[, .(count = sum(count)), by=.(seqnames, strand, TR_ID, sample)]
  
} else if(adapters == 'prime5') {
  TR.adapt.filt <- TR.adapt.count[correct_tss == T, .(count = sum(count)), by=.(seqnames, strand, TR_ID, sample)]
  
} else if(adapters == 'prime3') {
  TR.adapt.filt <- TR.adapt.count[correct_tes == T, .(count = sum(count)), by=.(seqnames, strand, TR_ID, sample)]
  
}
setnames(TR.adapt.filt, 'TR_ID', 'transcript_id')

###### Merge with read-transcript counts (possible filtered) #####
TR.gene.cluster.counts <- merge(TR.gene.cluster, 
                                TR.adapt.filt[,.(sample, transcript_id, count)],
                                by=c('transcript_id')
                                #all.x=T, 
                                #allow.cartesian=TRUE
                                )
```





### CAGE significance
```{r}
#TR.merged.data[,.N,CAGE_significance]


CAGE.sig <- fread("CAGEFighteR.SIG.tsv")


TR.merged.data[,CAGE_significance := NULL]


TR.merged.data <- merge(TR.merged.data[,], CAGE.sig[,.(CAGE_ID = gene, CAGE_significance)], by=c('CAGE_ID'), all.x=T)


TR.merged.data[,.N,CAGE_significance]
```


```{r}
#### summarise read-transcripts into ref transcripts
TR.ref.sum      <- TR.gene.cluster.counts[,.(read_count = sum(count)),  
  by=.(seqnames, gene_region, gene_cluster, gene, cmp_ref, Ref_Class, class_code, Kinetic_class, sample)]


## melt to get 0 counts
orf.counts <- dcast(TR.ref.sum, seqnames + gene_region + gene_cluster + gene + Kinetic_class + Ref_Class + class_code + cmp_ref ~ sample, value.var = 'read_count')
orf.perc   <- melt(orf.counts, variable.name = 'sample', value.name = 'read_count', id.vars = colnames(orf.counts)[1:8])
orf.perc[is.na(read_count), read_count := 0]

TR.ref.sum   <- orf.perc
## merge back for transcript strand
TR.ref.sum  <- merge(unique(TR.genes[,.(seqnames, strand, transcript_id)]), 
                     TR.ref.sum, by.y=c('seqnames', 'cmp_ref'), by.x=c('seqnames', 'transcript_id'))
setnames(TR.ref.sum, 'transcript_id', 'cmp_ref')
## and metadata
TR.ref.sum  <- merge(TR.ref.sum, metafilt[,metacols], by=c('sample'), all.x=T)



## Factorize Kinetic classes
TR.ref.sum[is.na(Kinetic_class), Kinetic_class := 'unknown']
TR.ref.sum[,Kinetic_class := factor(Kinetic_class, levels = c('IE', 'IE/E', 'E', 'IE/E/L', 'E/L', 'L', 'unknown'))]


```



```{r, eval=F}
#### Keep spliced genes only
spliced.genes <-  unique(TR.ref[exon_number > 1, gene_id])

message('Keeping spliced genes only!')
message('genes, other than: \n', paste(spliced.genes, collapse = '\n'), ' were filtered out!' )

## All Transcripts of Spliced Genes
spliced.TRs <- unique(viral.mrna[gene_id %in% spliced.genes, transcript_id]) 

TR.ref.sum  <- TR.ref.sum[ gene %in% spliced.genes, ]

```



```{r, eval=T}
#### Completely omit all non-validated isoform!
keep_equal_only <- T
class.code.tofilt <- '='
if(keep_equal_only) {
  TR.ref.sum <- TR.ref.sum[class_code %in% class.code.tofilt]
  message('Only the Reference Isoforms were kept, 
          all reads that are notequal ("=") to these, were omitted!')
}
```



```{r, eval=T}
#### Use the "gene" from the annotation, not the counting algorithm 
use_gene_from_annot <- T

if(use_gene_from_annot) {
  TR.ref.sum[,gene         := NULL]
  TR.ref.sum[,gene_cluster := NULL]
  
  TR.ref.sum.m <- merge(TR.ref.sum[],
                      TR.genes[,.(gene = gene_id, seqnames, strand, transcript_id)],
                      by.x=c('seqnames', 'strand', 'cmp_ref'),
                      by.y=c('seqnames', 'strand', 'transcript_id'),
                      all=F )
  TR.ref.sum.m <- merge(TR.ref.sum.m[],
                        unique(gene.clusters.all[, .(seqnames, strand, gene, gene_cluster)]),
                        by=c('seqnames', 'strand', 'gene'), all.x=T)
  if(nrow(TR.ref.sum.m) == nrow(TR.ref.sum)) { 
    message('The genes were assigned to the transcripts based on the Reference annotation!')
    TR.ref.sum <- TR.ref.sum.m
    rm(TR.ref.sum.m)
  } else {stop ()}
}
```

## Correct some genes in the annotation
```{r, eval=T}
correct_genes_with <- "NC_001491.2.mRNA_correct_genes_with.gff3"
correct_genes_with <- data.table(as.data.frame(rtracklayer::import.gff(correct_genes_with, version='3')))
#if(nrow(viral.ref[ID != transcript_id]) != 0 ) { stop('The IDs do not match for these transcripts: ', viral.ref[ID != transcript_id])}
correct_genes_with[type=='exon', exon_number := 1:.N, by = .(transcript_id)]
correct_genes_with <- correct_genes_with[type == 'mRNA', .(seqnames, source, type, phase, strand, start, end, transcript_id, Name, ID, gene_id, ORF_id)]


TR.ref.sum.corr <- merge(TR.ref.sum, 
                           correct_genes_with[,.(transcript_id, gene_id)],
                           by.x = c('cmp_ref'),
                           by.y = c('transcript_id'), 
                           all.x=T)

TR.corr <- unique(TR.ref.sum.corr[gene != gene_id, cmp_ref])

TR.ref.sum.corr <- TR.ref.sum.corr[gene != gene_id, ]
TR.ref.sum.corr[,gene := gene_id]
TR.ref.sum.corr[,gene_id := NULL]


TR.ref.sum <- rbind(
  TR.ref.sum.corr,
  TR.ref.sum[!cmp_ref %in% TR.corr, ]
)
```


```{r, eval=F}

TR.merged.data.corr <- merge(TR.merged.data, 
                           correct_genes_with[,.(transcript_id, gene_id)],
                           by.x = c('transcript_id'),
                           by.y = c('transcript_id'), 
                           all.x=T)

TR.corr <- unique(TR.merged.data.corr[gene != gene_id, transcript_id])

TR.ref.sum.corr <- TR.ref.sum.corr[gene != gene_id, ]
TR.ref.sum.corr[,gene := gene_id]
TR.ref.sum.corr[,gene_id := NULL]


TR.ref.sum <- rbind(
  TR.ref.sum.corr,
  TR.ref.sum[!cmp_ref %in% TR.corr, ]
)


```



DO NOT Filter second copies of genes, as in some cases there are transcripts from only one of the copies
```{r, eval=T}
filter_second_copy_genes <- F

if(filter_second_copy_genes) {
  ## Filter second copies of genes and rename first copy
  genes_to_filter <- grep('_2', unique(TR.ref.sum[, gene]), value = T)
  if(exists('genes_to_filter')) {
    message('The following genes were filtered out for counting: \n', paste(genes_to_filter, collapse = '\n'))
    TR.ref.sum <- TR.ref.sum[ ! gene %in% c(genes_to_filter), ]
  } else (message('No genes were filtered out for counting!'))
  
  if (all(
    gsub('_.*', '', grep('_', unique(TR.ref.sum[ , gene]), value = T)) %in%
    gsub('_.*', '', genes_to_filter)
  )) {
    
    TR.ref.sum[, gene         := gsub('_.*', '', gene)]
    TR.ref.sum[, gene_region  := gsub('_.*', '', gene_region)]
    TR.ref.sum[, gene_cluster := gsub('_.*', '', gene_cluster)]
    
    gene_regions <- unique(gsub('_.*', '', gene_regions))
    
    gene.plotdata <- data.frame(data.table(gene.plotdata)[!gene %in% genes_to_filter, ])
    #gene.plotdata[, gene         := gsub('_.*', '', gene)]
    #gene.plotdata[, gene         := gsub('_.*', '', gene)]
    #gene.plotdata <- unique(gsub('_.*', '', gene_regions))
    
  }  
} else {
  
  ## check for genes with two copies
  genes_to_filter <- grep('_2|_1', unique(TR.ref.sum[, gene]), value = T)
  
  ## assign Parent to them
  TR.ref.sum[gene %in% genes_to_filter, parent := gsub('_.*', '', gene)]
  ## we need to remove gene cluster as its not always the same
  TR.ref.sum[gene %in% genes_to_filter, gene_cluster := gsub('_.*', '', gene)]
  TR.ref.sum[gene %in% genes_to_filter, gene_region  := gsub('_.*', '', gene)]

  ## and overwrite gene as well
  TR.ref.sum[gene %in% genes_to_filter, gene := gsub('_.*', '', gene)]
  
  
    
}



  
#unique(TR.ref.sum[, gene_cluster])
```


```{r, eval=F}
### Filtering of very-low coverage samples
samples_to_filt <- c('C6_2h_3', 'PC-12_8h_3')


metafilt <- metafilt[!metafilt$sample %in% samples_to_filt, ]

TR.ref.sum <- TR.ref.sum[!sample %in% samples_to_filt, ]

message(paste(samples_to_filt, ' was filtered out!', collapse = '\n') )

```


###

```{r, eval=F}
### Torma and Fulop annotation and check

TX_nov_gff <- fread('EHV_transcript.tsv')
#TX_nov_gff[,All_Attributes := paste(c(paste(c(Attributes, TX_category), collapse = ';TX_Type='), gene_id, ORF_id, Name, ID), collapse=';'), by=.(ID)]
#TX_nov_gff <- TX_nov_gff[,.(Contig, Source, Type, `Start`, `End`, Score, Strand, Phase, All_Attributes)]
TX_nov_gff[,Attributes := NULL]
colnames(TX_nov_gff)[1:8]  <- c('seqnames', 'source', 'type', 'start', 'end', 'score', 'strand', 'phase')

rtracklayer::export.gff3(as.data.frame(TX_nov_gff), 'EHV_transcript.gff3')
```


```{r}
TX_nov_gff <- as.data.table(as.data.frame(rtracklayer::import.gff3('EHV_transcript.gff3')))
TX_nov_gff[, final_TX_ID := paste0('novTX_', final_TX_ID)]


TX_nov_gff[type == 'transcript', .N, by=TX_category]
TX_nov_gff[,.(width=abs(end-start)+1), by=TX_category][order(width)]

final_tx_nov_clust <- fread('final_tx_nov_clust.tsv', na.strings = '')
final_tx_nov_clust[, final_TX_ID := paste0('novTX_', final_TX_ID)]
```


##
```{r}
### merge the new transcripts into the original and omit those in the ori that were used to generate them

#sum(!trids.spliced %in% trids) ## OK

TR.merged.data.nov <- TX_nov_gff#[,.(transcript_id = gene=gene_id, )]
TR.merged.data.nov[,":="(transcript_id=final_TX_ID, class_code=TX_category,
                              gene=gene_id, gene_cluster=gene_id, gene_region=gene_id, 
                              exon_number=1, start.TR=start, end.TR=end, start.exon=start, end.exon=end, last_exon=T
                              )]
TR.merged.data.nov[,Ref_Class := paste(gene, class_code, final_TX_ID, sep = '-'), by=transcript_id]
TR.merged.data.nov[,":="(prime5.TR = fifelse(strand =='+', start, end), 
                              prime5.ex = fifelse(strand =='+', start, end), 
                              prime3.TR = fifelse(strand =='+', end, start), 
                              prime3.ex = fifelse(strand =='+', end, start)) ]
# OK 
## merge with CAGE

TR.merged.data.nov <- merge(TR.merged.data.nov, 
                                 unique(final_tx_nov_clust[,.(final_TX_ID=as.character(final_TX_ID), CAGE_ID, CAGE_significance, CAGE.cluster.start, CAGE.cluster.end, support)]),
                                 by='final_TX_ID', all.x=T )

TR.merged.data.nov[,thick.start := NA][, thick.end := NA]


TR.merged.data.nov[,transcript_id := Ref_Class]
```


### Add TX category
```{r}
TR.merged.data[ , TX_category := NULL]

TR.merged.data[ grepl('\\.[1-9]', transcript_id) &
               !grepl('PC', transcript_id) &
               !grepl('CX', transcript_id)   , TX_category := 'Putative mRNA']

unique(TR.merged.data[TX_category == 'Putative mRNA', transcript_id])

```


```{r}
cnames <- colnames(TR.merged.data)

TR.merged.data.nov <- unique(TR.merged.data.nov[,..cnames])

TR.merged.data <- rbind(TR.merged.data, 
                        TR.merged.data.nov[,..cnames], fill=F)


```



##
```{r}
TX_nov_gff <- as.data.table(as.data.frame(rtracklayer::import.gff3('EHV_transcript.gff3')))
TX_nov_gff[, final_TX_ID := paste0('novTX_', final_TX_ID)]

### merge the new transcripts into the original and omit those in the ori that were used to generate them

TR_IDs_nov    <- unique(final_tx_nov_clust[,.(transcript_id, TX_ID, TX_nov_ID, final_TX_ID)])
```



### DISTANCE <= 2
```{r, eval=F}
### merge the new transcripts into the original and omit those in the ori that were used to generate them

TR_IDs_nov    <- unique(final_tx_nov_clust[,
                      .(prime5.dist.ref = abs(prime5.TR - cluster_peak)), 
                      by=.(transcript_id, TX_ID, TX_nov_ID, final_TX_ID, prime5.TR, cluster_peak, TES)])


TR_IDs_nov    <- TR_IDs_nov[prime5.dist.ref <= 2]


TR_IDs_nov    <- unique(TR_IDs_nov[,.(transcript_id, TX_ID, TX_nov_ID, final_TX_ID)])

#length(unique(final_tx_nov_clust$cluster_peak))
#length(unique(final_tx_nov_clust$final_TX_ID))
#length(unique(final_tx_nov_clust$TES))

#unique(final_tx_nov_clust[,.(cluster_peak, final_TX_ID, TES)])
```


```{r}
TR_counts_nov <- merge(TR.adapt.filt, TR_IDs_nov, by='transcript_id')


TR_counts_final <- TR_counts_nov[,.(read_count = sum(count)), by=.(seqnames, strand, final_TX_ID, sample)]

#sum(!trids.spliced %in% trids) ## OK
#TR.ref.sum[Ref_Class == cmp_ref]

TR.ref.sum.nov <- TX_nov_gff#[,.(transcript_id = gene=gene_id, )]
TR.ref.sum.nov[,":="(final_TX_ID=final_TX_ID, class_code=TX_category, parent=NA, Kinetic_class = 'unknown',
                     gene=gene_id, gene_cluster=gene_id, gene_region=gene_id)]

TR.ref.sum.nov[,Ref_Class := paste(gene, class_code, final_TX_ID, sep = '-'), by=final_TX_ID]
TR.ref.sum.nov[,cmp_ref   := Ref_Class]

TR.ref.sum.nov <- TR.ref.sum.nov[,.(final_TX_ID, cmp_ref, seqnames, strand, gene, gene_region, Kinetic_class, Ref_Class, class_code, gene_cluster, parent)]

# OK 

TR.ref.sum.nov <- merge(TR.ref.sum.nov, TR_counts_final, by=c('seqnames', 'strand', 'final_TX_ID'))
TR.ref.sum.nov[,final_TX_ID := NULL]
```




```{r}
## melt to get 0 counts
orf.counts.nov <- dcast(TR.ref.sum.nov, 
                        seqnames + strand + gene_region + gene_cluster + gene + Kinetic_class + Ref_Class + 
                        class_code + cmp_ref + parent ~ sample, 
                        value.var = 'read_count')

orf.perc.nov   <- melt(orf.counts.nov, variable.name = 'sample', value.name = 'read_count', id.vars = c(1:10))
orf.perc.nov[is.na(read_count), read_count := 0]

TR.ref.sum.nov   <- orf.perc.nov

## and metadata
TR.ref.sum.nov  <- merge(TR.ref.sum.nov, metafilt[,metacols], by=c('sample'), all.x=T)


```





```{r}
TR.ref.sum <- rbind(TR.ref.sum, 
                    TR.ref.sum.nov, fill=F)


```


### Include NAGATA results
```{r}
source('NAGATA.R')

## window for maximum distance

nagata.mrna.TSS <- nagata.mrna[, .(seqnames, strand, start, end, transcript_id = ID, score, 
                                   prime5.TR = fifelse(strand == '+', start, end),
                                   prime3.TR = fifelse(strand == '-', start, end)) ]
nagata.mrna.TSS[, ":=" (start.TR  = start, end.TR = end,
                        start     = prime5.TR - 50, 
                        end       = prime5.TR + 50)]


TR.TSS <- unique(TR.merged.data[, .(seqnames, strand, start, end, transcript_id, source, TX_category,
                                    CAGE_significance, 
                                   prime5.TR ,
                                   prime3.TR ) ])


TR.TSS[, ":=" (start.TR  = start, end.TR = end,
               start     = prime5.TR - 50, 
               end       = prime5.TR + 50)]



nagata.ov <- foverlaps2(nagata.mrna.TSS, TR.TSS, by = c('seqnames', 'strand', 'start', 'end'), minoverlap=0, maxgap=0)

colnames(nagata.ov) <- gsub('i\\.', 'NAGATA.', colnames(nagata.ov))

nagata.ov[,distance := abs(prime5.TR - NAGATA.prime5.TR)]

nagata.ov[,min_dist := min(distance), by=.(seqnames, strand, transcript_id)]

nagata.ov <- nagata.ov[min_dist == distance,]

## Remove duplicated, because of a NAGATA transcript with same TSS but different TES
dup(nagata.ov$transcript_id)

nagata.ov <- nagata.ov[!duplicated(transcript_id),]
```


```{r}
nagata.stats <- nagata.ov[,.N, by=distance][order(N, decreasing = T)]

distance.mod <- nagata.stats[N == max(N), distance]


## number of NAGATA transcripts, not found in our annotation with distance <= 100
sum(!nagata.mrna.TSS$transcript_id %in% nagata.ov$NAGATA.transcript_id)

## Number of transcripts, supported by NAGATA from each source

nagata.ov[,.N, by=.(source)][order(N, decreasing = T)]

## Number of transcripts, supported by NAGATA from each source AND TX category
nagata.ov[,.N, by=.(source, TX_category)][order(N, decreasing = T)]


## Include all transcripts
nagata.ov.all <- merge(nagata.ov[,.(seqnames, strand, transcript_id, prime5.TR, NAGATA.transcript_id, NAGATA.prime5.TR, distance)], 
                         TR.TSS[,.(seqnames, strand, transcript_id, prime5.TR, TX_category, source, CAGE_significance)], 
                         by=c('seqnames', 'strand', 'transcript_id', 'prime5.TR'), all=T)
nagata.ov.all[is.na(source), source := 'NAGATA', ]

nagata.ov.all[is.na(distance), NAGATA.support := "NA"]
nagata.ov.all[distance <= 100, NAGATA.support := "distance <= 100"]
nagata.ov.all[distance <= 50,  NAGATA.support := "distance <= 50"]
nagata.ov.all[distance <= 25,  NAGATA.support := "distance <= 25"]

nagata.TX.stats <- nagata.ov.all[,.N, by=.(source, TX_category, NAGATA.support, CAGE_significance)][order(N, decreasing = T)]

```



```{r}

### Distribution of distances

gg.nt <- ggplot(nagata.ov, #[type == 'Putative'], 
                aes(distance)) +
          #geom_histogram(bins=100)+
          theme_ipsum() +
          geom_density(color='black') + # geom_density(aes(color=type)) +
          geom_vline(xintercept = distance.mod, color='blue') +
          geom_label(aes(distance.mod, 0.05), label=distance.mod, color='blue') +
          geom_vline(xintercept = 25, color='red') +
          geom_label(aes(25, 0.05), label=25, color='red')


ggsave('NAGATA-EHV transcripts TSS distance density.jpg', gg.nt, width = 10, height = 6)


gg.nt <- ggplot(nagata.ov, #[type == 'Putative'], 
                aes(distance)) +
          #geom_histogram(bins=100)+
          theme_ipsum() +
          geom_density(aes(color=source)) + # geom_density(aes(color=type)) +
          geom_vline(xintercept = distance.mod, color='blue') +
          geom_label(aes(distance.mod, 0.05), label=distance.mod, color='blue') +
          geom_vline(xintercept = 25, color='red') +
          geom_label(aes(25, 0.05), label=25, color='red')


ggsave('NAGATA-EHV transcripts TSS distance density BY SOURCE.jpg', gg.nt, width = 10, height = 6)


gg.nt <- ggplot(nagata.ov, #[type == 'Putative'], 
                aes(distance)) +
          #geom_histogram(bins=100)+
          theme_ipsum() +
          geom_density(aes(color=TX_category)) + # geom_density(aes(color=type)) +
          geom_vline(xintercept = distance.mod, color='blue') +
          geom_label(aes(distance.mod, 0.05), label=distance.mod, color='blue') +
          geom_vline(xintercept = 25, color='red') +
          geom_label(aes(25, 0.05), label=25, color='red')


ggsave('NAGATA-EHV transcripts TSS distance density BY TX-CATEGORY.jpg', gg.nt, width = 10, height = 6)


### Correlation of distance and NAGATA score

gg.ds <- ggplot(nagata.ov) +
          theme_ipsum() +
          geom_point(aes(distance, score, color=TX_category))

ggsave('NAGATA-EHV transcripts TSS distance vs NAGATA score.jpg', gg.ds, width = 10, height = 6)


### 
gg.sc <- ggplot(nagata.ov, aes(score)) +
          #geom_histogram(bins=100)+
          theme_ipsum() +
          geom_density(aes(color=TX_category))
          #geom_vline(xintercept = distance.mod, color='blue') +
          #geom_label(aes(distance.mod, 0.05), label=distance.mod, color='blue') +
          #geom_vline(xintercept = 25, color='red') +
          #geom_label(aes(25, 0.05), label=25, color='red')


```


```{r}
### Filter for 25 nt
nagata.ov    <- nagata.ov[distance <= 25,]

nagata.stats <- nagata.ov[,.N, by=distance][order(N, decreasing = T)]

distance.mod <- nagata.stats[N == max(N), distance]


## number of NAGATA transcripts, not found in our annotation with distance <= 100
sum(!nagata.mrna.TSS$transcript_id %in% nagata.ov$NAGATA.transcript_id)

## Number of transcripts, supported by NAGATA from each source

nagata.ov[,.N, by=.(source)][order(N, decreasing = T)]

## Number of transcripts, supported by NAGATA from each source AND TX category
nagata.ov[,.N, by=.(source, TX_category)][order(N, decreasing = T)]

```


```{r}
gg.nt <- ggplot(nagata.ov, aes(distance)) +
          #geom_histogram(bins=100)+
          theme_ipsum() +
          geom_density(color='black') +
          geom_vline(xintercept = distance.mod, color='blue') +
          geom_label(aes(distance.mod, 0.05), label=distance.mod, color='blue') +
          geom_vline(xintercept = 25, color='red') +
          geom_label(aes(25, 0.05), label=25, color='red')



ggsave('NAGATA-EHV transcripts TSS distance density.jpg', gg.nt, width = 10, height = 6)
```




### 
```{r}
nagata.ov.nov    <- nagata.ov[,]

nagata.ov.nov[, final_TX_ID := gsub('.*novTX_', 'novTX_', transcript_id)]


TX_nov_nagata_gff <- merge(TX_nov_gff[, .(seqnames, source, type, start, end, strand, final_TX_ID )], 
                        nagata.ov.nov[, .(seqnames, strand, score, final_TX_ID, distance, NAGATA.transcript_id, NAGATA.prime5.TR, NAGATA.prime3.TR )], by=c('seqnames', 'strand', 'final_TX_ID'), all.x=T )


rtracklayer::export.gff3(as.data.frame(TX_nov_nagata_gff), 'EHV_novel_NAGATA_merged.gff3')

```


### Merge NAGATA results
```{r}
TR.merged.data <- merge(TR.merged.data,
                        nagata.ov.all[,.(seqnames, strand, transcript_id, prime5.TR, NAGATA.transcript_id, NAGATA.prime5.TR, NAGATA.support)], 
                        by=c('seqnames', 'strand', 'transcript_id', 'prime5.TR'), all.x=T)


```



### Export gff
```{r}
TR.merged.data.EX  <- unique(
  TR.merged.data[,.(seqnames, strand, source, start=start.exon, end=end.exon, type = 'exon', ID=transcript_id, Name=transcript_id, TX_category)] )

TR.merged.data.TR  <- unique(
  TR.merged.data[,.(seqnames, strand, source, start=start.TR, end=end.TR, type = 'transcript', ID=transcript_id, Name=transcript_id, TX_category)] )

TR.merged.data.gff <- rbind(
  TR.merged.data.EX,
  TR.merged.data.TR  )


rtracklayer::export.gff3(TR.merged.data.gff, 'TR.merged.gff')

```





### Merge annotation and counts
```{r}
TR.merged.data.sp  <- 
    unique( TR.merged.data[,.(seqnames, strand, source, transcript_id, TX_category, 
                              NAGATA.transcript_id, NAGATA.prime5.TR, NAGATA.support, NAGATE.distance = abs(NAGATA.prime5.TR - prime5.TR),
                              CAGE_significance, CAGE_support=support, CAGE_score=score,
                              exon_number, exon_pos=paste0(start.exon, '-', end.exon)), by=transcript_id] )

TR.merged.data.sp  <- dcast(TR.merged.data.sp, ... ~ exon_number, 
                            #seqnames + strand + source + transcript_id + TX_category ~ exon_number, 
                            value.var = 'exon_pos')

setnames(TR.merged.data.sp, c('1', "2", "3", '4'), c('EX_1', "EX_2", "EX_3", 'EX_4'))


stopifnot(TR.ref.sum$Ref_Class %in% TR.merged.data.sp$transcript_id)
TR.merged.data.sp$transcript_id[!TR.merged.data.sp$transcript_id %in% TR.ref.sum$Ref_Class]


TR.ref.sum.sp  <- TR.ref.sum[,.(read_count = sum(read_count)), by=.(seqnames, strand, gene, hpi, Ref_Class) ]

#TR.ref.sum.sp[,presence := fifelse(read_count > 0, T, F), by=.(seqnames, strand, gene, hpi, Ref_Class)]

TR.ref.sum.sp  <- dcast(TR.ref.sum.sp, ... ~ hpi, value.var = 'read_count', fill=0)
TR.ref.sum.sp[,sum_dcDNA_read_count := `1h` + `2h` + `4h` + `6h` + `8h` + `12h` + `18h` + `24h` + `48h`, by=Ref_Class]
TR.ref.sum.sp$presence <- apply( TR.ref.sum.sp[,.(`1h`, `2h`, `4h`, `6h`, `8h`, `12h`, `18h`, `24h`, `48h`)], 1, function(x) length(x[x>0]) )
  


TR.all.data.sp  <- merge(TR.merged.data.sp, 
                            TR.ref.sum.sp,
                            by.x=c('seqnames', 'strand', 'transcript_id'), 
                            by.y=c('seqnames', 'strand', 'Ref_Class'), all=T)
```
### Include canonic annotation

```{r}

Torma_et_al <- fread('Torma_et_al_transcript.tsv')

Torma_canonic <- Torma_et_al[Canonic == '*' & `Transcript category` == 'Monocistronic', .(Gene, `Transcript ID`, Canonic)]


```


### Include CAGE canonic-to-truncated reatios

```{r}
## read in CAAGE 5-prime counts
CAGE.prime5.counts <- fread(paste0('CAGE/prime5.counts.tsv'), na.strings = '')
#prime3.counts <- fread(paste0('CAGE/prime3.counts.tsv'), na.strings = '')

## add window
CAGE.prime5.counts.sum <- CAGE.prime5.counts[,.(count=sum(count)), by=.(seqnames, strand, prime5, sample)]
CAGE.prime5.counts.sum[,':=' (start=prime5 - 10, end=prime5 + 10), by=.(seqnames, strand, prime5)]


## Transcript annotation
TR.CAGE.merge <- TR.merged.data[,.(start=prime5.TR - 10, end=prime5.TR + 10), by=.(seqnames, strand, prime5.TR, transcript_id, TX_category,  CAGE_ID)]

## add genes and transcript category info for canonics
TR.CAGE.merge <- unique(merge(Torma_canonic[,.(transcript_id=`Transcript ID`, Gene, Canonic)], TR.CAGE.merge, by='transcript_id', all.y=T))

TR.CAGE.merge <- merge(unique(TR.ref.sum[,.(transcript_id=Ref_Class, gene)]), TR.CAGE.merge, by='transcript_id', all.y=T)
TR.CAGE.merge <- TR.CAGE.merge[!is.na(gene)]


## only putatives and canonics
TR.CAGE.merge <- TR.CAGE.merge[ Canonic == '*' |
                                TX_category == 'Putative mRNA'
                               ,]

TR.CAGE.merge[is.na(TX_category) & transcript_id %in% Torma_canonic$`Transcript ID`, TX_category := 'Canonic']
TR.CAGE.merge[,.N,TX_category]

#TR.CAGE.merge[is.na(transcript_cat), transcript_cat := TX_category]
#TR.CAGE.merge[is.na(transcript_cat)]


## overlap it with transcript annotation
TR.CAGE.ov <- foverlaps2(CAGE.prime5.counts.sum,
                         TR.CAGE.merge, by=c('seqnames', 'strand', 'start', 'end'), minoverlap = 0, maxgap = 0)


colnames(TR.CAGE.ov) <- gsub('i\\.', 'CAGE.', colnames(TR.CAGE.ov))

TR.CAGE.ov[,distance := abs(prime5.TR - prime5)]

## find the transcript that was closer to the CAGE signal
TR.CAGE.ov[,min_dist := min(distance), by=.(seqnames, strand, prime5, sample)]

TR.CAGE.ov <- TR.CAGE.ov[min_dist == distance,]

## in case where a CAGE signal is exactly in the middle of two TSSs, assign it to the longer one
TR.CAGE.ov[,longer.prime5 := fifelse(strand == '-', max(prime5.TR), min(prime5.TR)), by=.(seqnames, strand, prime5, sample)]
TR.CAGE.ov <- TR.CAGE.ov[prime5.TR == longer.prime5, ]

## split the counts into equal number if there are still duplicates
TR.CAGE.ov[,n_TR := n_distinct(transcript_id), by=.(seqnames, strand, prime5, sample)]
TR.CAGE.ov[n_TR > 1, count := count / n_TR]

TR.CAGE.ov[, n_strand := n_distinct(strand), by=.(seqnames, prime5, sample)]

#TR.CAGE.ov.dup <- unique(TR.CAGE.ov[duplicated(prime5) & n_strand == 1, transcript_id])
#TR.CAGE.ov[transcript_id %in% TR.CAGE.ov.dup]
```


```{r, fig.height=12, fig.width=7}
## count the prime5 ends 
TR.CAGE.ov.prime5.counts <- TR.CAGE.ov[,.(TX_count = sum(count)), by=.(seqnames, strand, sample, transcript_id, TX_category, gene, Canonic, CAGE_ID, prime5.TR)]

## make ratio
TR.CAGE.ov.prime5.counts[, gene_sum_count := sum(TX_count), by=.(seqnames, strand, sample, gene)]
##
TR.CAGE.ov.prime5.counts[, TX_ratio := TX_count / gene_sum_count]


TR.CAGE.ov.prime5.ratio <- merge(
  unique(TR.CAGE.ov.prime5.counts[TX_category == 'Canonic', .(seqnames, strand, gene, sample, Canonic_ID=transcript_id, Canonic_count=TX_count)]),
  unique(TR.CAGE.ov.prime5.counts[TX_category != 'Canonic', .(seqnames, strand, gene, sample, Isoform_ID=transcript_id, Isoform_count=TX_count)]),
  by=c('seqnames', 'strand', 'gene', 'sample'))

TR.CAGE.ov.prime5.ratio[, Isoform_ratio := Isoform_count / Canonic_count, by=.(seqnames, strand, gene, sample)]

TR.CAGE.ov.prime5.ratio[, Isoform_ratio_mean := mean(Isoform_ratio), by=.(seqnames, strand, gene)]

#TR.CAGE.ov.prime5.ratio <- dcast(TR.CAGE.ov.prime5.counts, seqnames + strand + gene + sample ~ TX_category , value.var = 'TX_ratio')
#TR.CAGE.ov.prime5.ratio <- dcast(TR.CAGE.ov.prime5.counts, seqnames + strand + transcript_id + TX_category ~ sample, value.var = 'TX_ratio')


ggplot(TR.CAGE.ov.prime5.ratio, aes(Isoform_ratio)) + 
  geom_histogram(bins=100) +
  theme_ipsum() + 
  facet_wrap(~sample, ncol=1)
```


```{r, fig.height=12, fig.width=7}

TR.CAGE.ov.prime5.ratio.uni <- unique(TR.CAGE.ov.prime5.ratio[,
              .(seqnames, strand, gene, Canonic_ID, Isoform_ID, Isoform_ratio_mean)])


dup(TR.CAGE.ov.prime5.ratio.uni$Isoform_ID)


ggplot(TR.CAGE.ov.prime5.ratio.uni, aes(Isoform_ratio_mean)) + 
  geom_histogram(bins=100) +
  theme_ipsum() 

```


### Add CAGE Isoform ratio to tr.all.data
```{r}

#### Minimum CAGE-TSS ratio
CAGE_TSS_ratio_thresh  <- 0.05


TR.all.data.sp <- merge(TR.all.data.sp, 
                        TR.CAGE.ov.prime5.ratio.uni[,.(seqnames, strand, gene, transcript_id=Isoform_ID, CAGE_TSS_ratio=Isoform_ratio_mean)],
                        by=c('seqnames', 'strand', 'gene', 'transcript_id'), all.x=T)

```


### Add categories from Torma

```{r}

Torma_categories <- fread('TR.all.data.sp.Torma.tsv')

Torma_categories <- Torma_categories[,.(transcript_id, TX_category2=TX_category)]

TR.all.data.spx  <- merge(Torma_categories, TR.all.data.sp, by=c('transcript_id'), all=T)

if (nrow(TR.all.data.spx) == nrow(TR.all.data.sp)) {
  
  TR.all.data.spx[,TX_category  := TX_category2]
  TR.all.data.spx[,TX_category2 := NULL]
  
  TR.all.data.sp <- TR.all.data.spx
  
  rm(TR.all.data.spx)
  
}

TR.all.data.sp[,.N,TX_category]

```




```{r, fig.width=12, fig.height=10}
### all transcripts

gg1 <- ggVennDiagram::ggVennDiagram(
  list(
    dcDNA = TR.all.data.sp[presence >= 3 & sum_dcDNA_read_count >= 5 , transcript_id],
    CAGE  = TR.all.data.sp[CAGE_significance %in% c('**', '***') , transcript_id],
    dRNA  = TR.all.data.sp[NAGATA.support %in% c('distance <= 25') , transcript_id]
  )
)

gg2 <- ggVennDiagram::ggVennDiagram(
  list(
    dcDNA = TR.all.data.sp[presence >= 1 & sum_dcDNA_read_count >= 3 , transcript_id],
    CAGE  = TR.all.data.sp[CAGE_significance %in% c('**', '***') , transcript_id],
    dRNA  = TR.all.data.sp[NAGATA.support %in% c('distance <= 25') , transcript_id]
  )
)

gg3 <- ggVennDiagram::ggVennDiagram(
  list(
    dcDNA = TR.all.data.sp[presence >= 3 & sum_dcDNA_read_count >= 3, transcript_id],
    CAGE  = TR.all.data.sp[CAGE_significance %in% c('**', '***') , transcript_id],
    dRNA  = TR.all.data.sp[NAGATA.support %in% c('distance <= 25') , transcript_id]
  )
)

gg4 <- ggVennDiagram::ggVennDiagram(
  list(
    dcDNA = TR.all.data.sp[presence >= 3 & sum_dcDNA_read_count >= 3, transcript_id],
    CAGE  = TR.all.data.sp[CAGE_significance %in% c('**', '***') & TR.all.data.sp$CAGE_TSS_ratio >= CAGE_TSS_ratio_thresh, transcript_id],
    dRNA  = TR.all.data.sp[NAGATA.support %in% c('distance <= 25') , transcript_id]
  )
)


cowplot::plot_grid(gg1, gg2, gg3, gg4, nrow = 2) + ggtitle('all transcripts')

ggsave('All transcripts.venn.jpg', width=12, height=7)

```


```{r, fig.width=12, fig.height=10}
### Torma et al transcripts

gg1 <- ggVennDiagram::ggVennDiagram(
  list(
    dcDNA = TR.all.data.sp[presence >= 3 & sum_dcDNA_read_count >= 5 & source == 'Torma_et_al' , transcript_id],
    CAGE  = TR.all.data.sp[CAGE_significance %in% c('**', '***')     & source == 'Torma_et_al', transcript_id],
    dRNA  = TR.all.data.sp[NAGATA.support %in% c('distance <= 25')   & source == 'Torma_et_al', transcript_id]
  )
)

gg2 <- ggVennDiagram::ggVennDiagram(
  list(
    dcDNA = TR.all.data.sp[presence >= 1 & sum_dcDNA_read_count >= 3  & source == 'Torma_et_al', transcript_id],
    CAGE  = TR.all.data.sp[CAGE_significance %in% c('**', '***')      & source == 'Torma_et_al', transcript_id],
    dRNA  = TR.all.data.sp[NAGATA.support %in% c('distance <= 25')    & source == 'Torma_et_al', transcript_id]
  )
)

gg3 <- ggVennDiagram::ggVennDiagram(
  list(
    dcDNA = TR.all.data.sp[presence >= 3 & sum_dcDNA_read_count >= 3  & source == 'Torma_et_al', transcript_id],
    CAGE  = TR.all.data.sp[CAGE_significance %in% c('**', '***')      & source == 'Torma_et_al', transcript_id],
    dRNA  = TR.all.data.sp[NAGATA.support %in% c('distance <= 25')    & source == 'Torma_et_al', transcript_id]
  )
)

gg4 <- ggVennDiagram::ggVennDiagram(
  list(
    dcDNA = TR.all.data.sp[presence >= 3 & sum_dcDNA_read_count >= 3 & source == 'Torma_et_al', transcript_id],
    CAGE  = TR.all.data.sp[CAGE_significance %in% c('**', '***')     & source == 'Torma_et_al' & TR.all.data.sp$CAGE_TSS_ratio >= CAGE_TSS_ratio_thresh, transcript_id],
    dRNA  = TR.all.data.sp[NAGATA.support %in% c('distance <= 25')   & source == 'Torma_et_al', transcript_id]
  )
)


cowplot::plot_grid(gg1, gg2, gg3, gg4, nrow = 2) + ggtitle('all transcripts')

ggsave('Torma et al.venn.jpg', width=12, height=7)

```


```{r, fig.width=12, fig.height=10}

### Tomb치cz et al transcripts
gg1 <- ggVennDiagram::ggVennDiagram(
  list(
    dcDNA = TR.all.data.sp[presence >= 3 & sum_dcDNA_read_count >= 5 & source == 'Tombacz_et_al', transcript_id],
    CAGE  = TR.all.data.sp[CAGE_significance %in% c('**', '***')     & source == 'Tombacz_et_al', transcript_id],
    dRNA  = TR.all.data.sp[NAGATA.support %in% c('distance <= 25')   & source == 'Tombacz_et_al', transcript_id]
  )
)

gg2 <- ggVennDiagram::ggVennDiagram(
  list(
    dcDNA = TR.all.data.sp[presence >= 1 & sum_dcDNA_read_count >= 3 & source == 'Tombacz_et_al', transcript_id],
    CAGE  = TR.all.data.sp[CAGE_significance %in% c('**', '***')     & source == 'Tombacz_et_al', transcript_id],
    dRNA  = TR.all.data.sp[NAGATA.support %in% c('distance <= 25')   & source == 'Tombacz_et_al', transcript_id]
  )
)

gg3 <- ggVennDiagram::ggVennDiagram(
  list(
    dcDNA = TR.all.data.sp[presence >= 3 & sum_dcDNA_read_count >= 3 & source == 'Tombacz_et_al', transcript_id],
    CAGE  = TR.all.data.sp[CAGE_significance %in% c('**', '***')     & source == 'Tombacz_et_al', transcript_id],
    dRNA  = TR.all.data.sp[NAGATA.support %in% c('distance <= 25')   & source == 'Tombacz_et_al', transcript_id]
  )
)

gg4 <- ggVennDiagram::ggVennDiagram(
  list(
    dcDNA = TR.all.data.sp[presence >= 3 & sum_dcDNA_read_count >= 3 & source == 'Tombacz_et_al', transcript_id],
    CAGE  = TR.all.data.sp[CAGE_significance %in% c('**', '***')     & source == 'Tombacz_et_al' & TR.all.data.sp$CAGE_TSS_ratio >= CAGE_TSS_ratio_thresh, transcript_id],
    dRNA  = TR.all.data.sp[NAGATA.support %in% c('distance <= 25')   & source == 'Tombacz_et_al', transcript_id]
  )
)

cowplot::plot_grid(gg1, gg2, gg3, gg4, nrow = 2) + ggtitle('Tomb치cz et al transcripts')


ggsave('Tomb치cz et al transcripts.venn.jpg', width=12, height=7)
```


```{r, fig.width=12, fig.height=10}

### Putative mRNAs
gg1 <- ggVennDiagram::ggVennDiagram(
  list(
    dcDNA = TR.all.data.sp[presence >= 3 & sum_dcDNA_read_count >= 5 & source == 'Tombacz_et_al' & TX_category == 'Putative mRNA', transcript_id],
    CAGE  = TR.all.data.sp[CAGE_significance %in% c('*', '**', '***')     & source == 'Tombacz_et_al' & TX_category == 'Putative mRNA', transcript_id],
    dRNA  = TR.all.data.sp[NAGATA.support %in% c('distance <= 25')   & source == 'Tombacz_et_al' & TX_category == 'Putative mRNA', transcript_id]
  )
)

gg2 <- ggVennDiagram::ggVennDiagram(
  list(
    dcDNA = TR.all.data.sp[presence >= 1 & sum_dcDNA_read_count >= 3 & source == 'Tombacz_et_al'  & TX_category == 'Putative mRNA', transcript_id],
    CAGE  = TR.all.data.sp[CAGE_significance %in% c('*', '**', '***')     & source == 'Tombacz_et_al'  & TX_category == 'Putative mRNA', transcript_id],
    dRNA  = TR.all.data.sp[NAGATA.support %in% c('distance <= 25')   &  source == 'Tombacz_et_al' & TX_category == 'Putative mRNA', transcript_id]
  )
)

gg3 <- ggVennDiagram::ggVennDiagram(
  list(
    dcDNA = TR.all.data.sp[presence >= 3 & sum_dcDNA_read_count >= 3 & source == 'Tombacz_et_al' & TX_category == 'Putative mRNA', transcript_id],
    CAGE  = TR.all.data.sp[CAGE_significance %in% c('*', '**', '***')     & source == 'Tombacz_et_al' & TX_category == 'Putative mRNA', transcript_id],
    dRNA  = TR.all.data.sp[NAGATA.support %in% c('distance <= 25')   & source == 'Tombacz_et_al' & TX_category == 'Putative mRNA', transcript_id]
  )
)

gg4 <- ggVennDiagram::ggVennDiagram(
  list(
    dcDNA = TR.all.data.sp[presence >= 3 & sum_dcDNA_read_count >= 3 & source == 'Tombacz_et_al' & TX_category == 'Putative mRNA', transcript_id],
    CAGE  = TR.all.data.sp[CAGE_significance %in% c('*', '**', '***')     & source == 'Tombacz_et_al' & TX_category == 'Putative mRNA' & TR.all.data.sp$CAGE_TSS_ratio >= CAGE_TSS_ratio_thresh, transcript_id],
    dRNA  = TR.all.data.sp[NAGATA.support %in% c('distance <= 25')   & source == 'Tombacz_et_al' & TX_category == 'Putative mRNA', transcript_id]
  )
)

cowplot::plot_grid(gg1, gg2, gg3, gg4, nrow = 2) + ggtitle('Putative mRNAs')

ggsave('Putative mRNAs.venn.jpg', width=12, height=10)
```


### Filter transcripts based on dRNA-NAGATA, CAGE-TSS-ratio and dcDNA-presence
```{r}

putatives_to_keep <- unique(
    TR.all.data.sp[  presence >= 3 & sum_dcDNA_read_count >= 3
                   & CAGE_significance %in% c('*', '**', '***') & TR.all.data.sp$CAGE_TSS_ratio >= CAGE_TSS_ratio_thresh 
                   & NAGATA.support %in% c('distance <= 25')     
                   & source == 'Tombacz_et_al' & TX_category == 'Putative mRNA', transcript_id]
)

all_non_putative <-  unique(
    TR.all.data.sp[ TX_category != 'Putative mRNA' & !is.na(TX_category), transcript_id]
)

Torma_transcripts <- unique(
    TR.all.data.sp[ source == 'Torma_et_al', transcript_id]
)

TRs.to.keep <- unique(c(all_non_putative, putatives_to_keep, Torma_transcripts))


#TR.all.data.sp[,.N,TX_category]



### Filter

TR.ref.sum  <- TR.ref.sum[Ref_Class %in% TRs.to.keep]

TR.merged.data <- TR.merged.data[transcript_id %in% TRs.to.keep]


TR.all.data.sp[transcript_id %in% putatives_to_keep, putatives_to_keep := T]

```


### Export
```{r}

fwrite(TR.all.data.sp, 'TR.all.data.sp.tsv', sep='\t')

```

####





## START MAIN WF

```{r}

orf.sum  <- TR.ref.sum
orf.perc <- orf.sum


#orf.perc[grepl('1h', sample),.(sum_read_count=sum(read_count)), by=.(gene, sample)][sum_read_count != 0]
```





## Count normalization
```{r, eval=F}
## THE COUNTS ARE NORMALIZED TO THEIR RESPECTIVE PARENT GENE OR GENE/CLUSTER COUNTS AND NOT FOR LIBRARY SIZE OR OTHER !

### Normalize counts

#### Normalization method
norm.method <- 'average_genome'
multip      <- 10000

 if (norm.method == 'average_genome') {
  
  ## Average genome norm method
  norm_cov_summary <-  fread(paste0(outdir, '/norm.cov.summary.tsv'))
  normBase <- norm_cov_summary[,.(sample, average_coverage)]
  colnames(normBase)[2] <- 'norm_base'
  
  fig.dir <- paste0(fig.dir, '.norm_genome')

} else if (norm.method == 'viral_read.count') {
    
  ## Viral read count method
  # sum up all the ORF counts

  normBase <- melt(gene.sample_count.sp, variable.name = 'sample', value.name = 'count')
  normBase <- normBase[,.(norm_base = sum(count)), by=sample]
  
  fig.dir <- paste0(fig.dir, '.norm_LoRTIA')
  
} else if (norm.method == 'total_read.count') {
  ## For total read count method

  # This is not the LoRTIA-processed read counts, so each read is counted twice because they are from dcDNA that contains both orientations. 
  # But since this is true for both the virus and the host, this will cause each ORF to be estimated down approx 2-fold but in an unbiased way.

  normBase <- fread('../rebasecall/read_count.fastq.txt')
  normBase[ , sample := gsub('.fastq', '', file)]
  normBase[ , norm_base := read_count]
  normBase <- normBase[,.(sample, norm_base)]
  
  fig.dir <- paste0(fig.dir, '.norm_total')
  
} else if (norm.method == 'host_read.count') {
  ## For total HOST read count method

  # This is not the LoRTIA-processed read counts, so each read is counted twice because they are from dcDNA that contains both orientations. 
  # But since this is true for both the virus and the host, this will cause each ORF to be estimated down approx 2-fold but in an unbiased way.
  # Currently this is the unmapped read count!

  normBase <- fread('../rebasecall/read_count.txt')
  #normBase[ , sample := gsub('.fastq', '', file)]
  #normBase <- normBase[,.(sample, unmapped)]
  normBase[ , norm_base := unmapped]
  normBase <- normBase[,.(sample, norm_base)]
  
  fig.dir <- paste0(fig.dir, '.norm_host')
  
} else if (norm.method == 'WO') {
  
  ## NO normalization of counts

  normBase <- fread('../rebasecall/read_count.txt')
  normBase[ , norm_base := 1]
  normBase <- normBase[,.(sample, norm_base)]
  
  fig.dir <- paste0(fig.dir, '.norm_Without')
  
}

try({ dir.create(fig.dir) })

##
normBase <- normBase[order(normBase$sample, colnames(orf.counts)[-c(1:6)]), ]
stopifnot(all(normBase$sample == colnames(orf.counts)[-c(1:6)] ))

orf.perc <- merge(normBase, orf.sum, by='sample')
orf.perc[,read_count := read_count / norm_base]

```




# Counts based on iterative GFF-compare results  

This was done as follows:  

1. The reference transcripts were compared individually/iteratively to the reads.  
2. The distances between the reads and each reference transcript hit were calculated.
3. For each read, the most similar reference transcript was assigned (lowest distance).  
4. Additionally, these distances must had fulfill the following rules: 
*max prime5 distance <= 10*  
*max prime3 distance <= 10*  
*exon/intron junction difference <= 2*  

5. Thus based on these, the read counts for each reference transcript were assessed.
6. The transcripts were then assigned to their parent genes (by Torma G) and their ratio (isoform ratio) was calculated as follows:  
*isoform ratio = transcript isoform count / total gene count*  

7. Those reads that could not be assigned to the reference transcripts were assigned to their parent genes, based on their 5-prime ends and independently to their parent gene cluster, based on their 3-prime ends.  
So the reads are either:  
- assigned to a reference transcript
- could not not be assigned to a ref TR, in which case they were:  
  1. assigned to their parent gene based on their *5-prime* end overlapping with the gene's *TSS*  
  2. assigned to their parent gene-cluster based on their *3-prime* end overlapping with the cluster's *TES*  
  3. assigned to both  



## 1. Kinetics of Reference Transcript Isoforms, according to their host gene {.tabset}

The reads were assigned to their parent gene based on their *5-prime* end overlapping with the gene's *TSS*  
(but not necessarily with the gene cluster's TES)

```{r}
#### NO NORMALIZATION HERE
orf.perc[,norm_base := 1]

#### Keep only those reads, which could be assigned to a gene (based on '=' ref transcripts or 5-prime overlaps)
orf.sum <- orf.perc[!is.na(gene),]


#### Summarize the counts on Ref_Class and Gene
ref.sum      <- orf.sum[, 
                        .(read_count = sum(read_count)),
                        by=.(seqnames, gene, Ref_Class)]

#### Gene-level summary
orf.sum.gene.TX <- orf.sum[, 
                        .(read_count = sum(read_count)),
                        by=.(sample, ID=gene, seqnames, strand, 
                             gene_cluster=gene, gene, Kinetic_class, 
                             hpi, Time, cell_line, group)]

```


```{r}
#### Exclude Reference Isoforms with counts below the threshold
Ref.class.thresh <- 1
Ref.class.tofilt <- ref.sum[read_count < Ref.class.thresh, Ref_Class]

if(length(Ref.class.tofilt) > 0) {
  message('The following Reference Isoforms (n=', length(Ref.class.tofilt), ') had total counts below the threshold (', Ref.class.thresh, ')
across every sample, and thus were filtered out from the counting: \n', 
  paste(Ref.class.tofilt, collapse = '\n'))
  
  orf.sum <- orf.sum[!Ref_Class %in% Ref.class.tofilt]
  
} else (message('No Reference Isoforms were filtered out for counting!'))



#### Overwite Ref_class for simplicity (either Reference Isoform name or class code)
orf.sum[,Ref_Class := ifelse(class_code == '=', cmp_ref, class_code)]

TX.gene.uni <- unique(orf.sum[, .(seqnames,  strand, Ref_Class, 
                                  gene, gene_region, gene_cluster)])

```


```{r}
#### DO NOT !! Summarize the counts on Ref_Class and Gene
#### WHY NOT ???? 
#orf.sum      <- orf.sum[, .(read_count = sum(read_count)),
#                        by=.(seqnames, gene, Ref_Class, Kinetic_class, sample, norm_base, group, hpi, Time, cell_line)]
##
orf.sum.perc <- orf.sum


### Summarise counts to get mean count per gene
## sum counts per gene
orf.sum.perc[,sum_read_count  := sum(read_count),  
             by=.(seqnames, gene, norm_base, sample, group, hpi, Time, cell_line)]
## normalize counts per gene
orf.sum.perc[,norm_read_count := read_count / sum_read_count] 
orf.sum.perc[is.na(norm_read_count), norm_read_count := 0] 


## statistics of normalized counts
orf.perc.mean <- orf.sum.perc[,.(mean=mean(norm_read_count), sd=sd(norm_read_count), varcoeff=sd(norm_read_count) / mean(norm_read_count)), 
                             by=.(seqnames, gene, gene_cluster, gene_region, Ref_Class, Kinetic_class, group, hpi, Time, cell_line)]

## merge back summary with counts
orf.sum.perc <- merge(orf.sum.perc, orf.perc.mean, 
                      by=c('seqnames', 'Ref_Class', 'gene', 'gene_cluster', 'gene_region', 'Kinetic_class',  'hpi', 'Time', 'cell_line', 'group'))



## add a group ID to the isoforms
TR.genes.ID <- unique(orf.sum.perc[,.(gene, Ref_Class)])
TR.genes.ID[,Isoform_nr :=  seq(1,n_distinct(Ref_Class)), by=.(gene)]
TR.genes.ID[,Isoform_nr :=  as.factor(as.character(Isoform_nr))]

orf.sum.perc <- merge(orf.sum.perc, TR.genes.ID, by=c('gene', 'Ref_Class')) #, allow_cartesian=T)

## Merge with gff-compare class-code descriptions 

orf.sum.perc <- merge(orf.sum.perc, gff_compare_classes[,.(`Class Code`, Description = `One-word Description`)], by.x=c('Ref_Class'), by.y=c('Class Code'), all.x=T)

orf.sum.perc[ , Ref_Class := fifelse(is.na(Description), Ref_Class, paste0(Description, ' ("', Ref_Class, '")')) ]


## Add replicate nr.
orf.sum.perc[, rep := factor(gsub('.*h_', '', sample))]
orf.sum.perc[, hpi := factor(hpi, levels=as.character(levels(metafilt$hpi)))]

## Include novel combinations
#genes_and_clusters <- na.omit(unique(orf.sum.perc[, .(gene, gene_cluster)]))[, .(CompleteList = append_cluster(gene, gene_cluster)), by = gene_cluster]

# Combine all results into one vector and ensure uniqueness
#gene_regions  <- unique(c(gene_regions, unique(unlist(genes_and_clusters$CompleteList)) ))



## Factorize genes and clusters

gm  <-  setdiff(unique(na.omit(orf.sum.perc[,gene])), gene_regions)
gcm <-  setdiff(unique(na.omit(orf.sum.perc[,gene_cluster])), gene_regions)
grm <-  setdiff(unique(na.omit(orf.sum.perc[,gene_region])), gene_regions)
ga  <-  unique(c(gm, gcm, grm))

if(length(ga) != 0 ) {
  gene_regions <- c(ga, gene_regions)
  message(paste0(ga, collapse=' ;'), ' was (were) not in the gene list, so it (they) was (were) added!')
}

orf.sum.perc[,gene         := factor(gene,         levels=gene_regions)]
orf.sum.perc[,gene_cluster := factor(gene_cluster, levels=gene_regions)]
orf.sum.perc[,gene_region  := factor(gene_region,  levels=gene_regions)]

# sum read counts in 1h
#orf.perc[grepl('1h', sample),.(sum_read_count=sum(read_count)), by=.(gene, sample)][sum_read_count != 0]
#orf.sum.perc[grepl('1h', sample),.(sum_read_count=sum(read_count)), by=.(gene, sample)][sum_read_count != 0]
```


Merge annotation with RefClass
```{r}

TR.merged.data <- merge(unique(TR.ref.sum[,.(cmp_ref, Ref_Class, gene, gene_cluster, gene_region)]),
                        #unique(orf.sum.perc[,.(cmp_ref, Ref_Class, gene, gene_cluster, gene_region)]), 
                        TR.merged.data, 
                        by.x='cmp_ref', by.y='transcript_id', all.y=T)
colnames(TR.merged.data)[1] <- 'transcript_id'


TR.merged.data.gene <- data.table(as.data.frame(TR.merged.data))
```



Exclude those ref transcripts that were not observed (or observed under the threshold) from the annotation as well
```{r, eval=T}
## exclude those ref transcripts that were not observed (or observed under the threshold)
## from the annotation as well

TRs.to.plot <- unique(orf.sum.perc$cmp_ref)


TR.merged.data.gene <- TR.merged.data.gene[transcript_id %in% TRs.to.plot,]

```


```{r}

#TR.merged.data.gene[grep('novTX', transcript_id), Ref_Class := paste0(gene, '-')]
#orf.sum.perc[grep('novTX', cmp_ref), Ref_Class := paste0(Ref_Class, ' - this work')]

orf.sum.perc[grep('novTX', cmp_ref), Ref_Class := cmp_ref]

```

### Combine Transcript counts with annotation -->> Supp Table S2
```{r}

TR_counts_all <- dcast(
  unique(orf.sum.perc[,.(seqnames, strand, gene, Ref_Class, sample, read_count)]), 
  seqnames + strand + gene + Ref_Class ~ sample, value.var = 'read_count', fill = 0)


supp_tab_S2 <- merge(TR.merged.data.gene[,
      .(seqnames, gene, strand, Ref_Class, source,  start.TR, end.TR, final_TX_ID = gsub('.*novTX', 'novTX', Ref_Class),
        exon_number, start.exon, end.exon,  
        CAGE.cluster.start, CAGE.cluster.end, CAGE_ID, CAGE_significance, CAGE_score = score, CAGE_support=support)], 
      TX_nov_gff[, .(final_TX_ID, TX_category)], by='final_TX_ID', all.x=T)

supp_tab_S2 <- merge(supp_tab_S2, TR_counts_all, by=c('seqnames', 'strand', 'gene', 'Ref_Class'))

                     

```

```{r}
## Filter based on transcript ratio

min.IF.ratio <- 0.005

orf.sum <- orf.sum.perc
orf.sum[, sum_abundance  := sum(norm_read_count), by=.(Ref_Class, gene, hpi)]
orf.sum[, mean_abundance := max(norm_read_count), by=.(Ref_Class, gene, hpi)]
orf.sum[, min_abundance  := min(norm_read_count), by=.(Ref_Class, gene)]
orf.sum[, max_abundance  := max(norm_read_count), by=.(Ref_Class, gene)]


orf.sum <- unique(orf.sum[,.(Ref_Class, gene, hpi, mean_abundance, sum_abundance, min_abundance, max_abundance)])

trs.tofilt <- unique(orf.sum[max_abundance < min.IF.ratio, Ref_Class])


if(length(trs.tofilt) > 0) {
  message('The following Reference Isoforms (n=', length(trs.tofilt), ') had a minimum isoform ratio below the threshold (', min.IF.ratio*100, '%)
across every sample, and thus were filtered out from the counting: \n', 
  paste(trs.tofilt, collapse = '\n'))
  
} else (message('No Reference Isoforms were filtered out for counting!'))



orf.sum.perc        <- orf.sum.perc[!Ref_Class %in% trs.tofilt,]
TR.merged.data.gene <- TR.merged.data.gene[!Ref_Class %in% trs.tofilt,]


```


```{r}

mean_abundance <- dcast(
  unique(orf.sum.perc[,.(seqnames, strand, gene, Ref_Class, hpi, mean)]), 
  seqnames + strand + gene + Ref_Class ~ hpi, value.var = 'mean', fill = 0)


sd_abundance <- dcast(
  unique(orf.sum.perc[,.(seqnames, strand, gene, Ref_Class, hpi, sd)]), 
  seqnames + strand + gene + Ref_Class ~ hpi, value.var = 'sd', fill = 0)

```


# Plotting


### Annotation plotting function
```{r}
figw  <- 20
figh  <- 9

gene.sizes <- data.frame(gene_arrowhead_height=6,
                         gene_arrow_body_height=4,
                         gene_label_height=2.5,   ## size of gene label text
                         gene_feature_height=0, ## lineheight of feature line
                         gene_feature_width=5,
                         gene_feature_label_height=0, ## position of gene label text
                         gene_feature_label_text=-2, ## inactive
                         unstranded_rect_height=0.5,
                         unstranded_feature_height=6,
                         unstranded_feature_width=6,
                         unstranded_feature_label_height=5,
                         unstranded_feature_label_text=6,
                         genome_feature_arrowhead_height=2,
                         genome_feature_arrow_body_height=2,
                         ## introns
                         linetype='dashed', 
                         linewidth=0.01
)
angle <- 0


transcript.sizes <- data.frame(
  transcript.label.size = 2,
  exon.rect_height = 0.1, exon.rect.linewidth=0.01,  
  intron.linetype = 'dashed', intron.linewidth = 0.1)

sizes <- 6


scales <- 'fixed'

## scale of annotation to coverage
genomplot.scale <- 6

breakseq <- 2000

## plot size
width <- 35; height <- 14


ylims.gene <- NULL #c(-3.5, 0) ## 

gene_to_plot <- 'NOIR'

vis_win <- 500

ann.plotfun <- function(gene_to_plot=NULL, plot.data=plot.data, 
                        visfrom = NULL, visto = NULL, vis_win=500, 
                        TR.merged.data.gene = NULL, 
                        breakseq=5000,
                        ...) {
  
  if(!is.null(gene_to_plot) & is.null(visfrom) & is.null(visto)) {
    visfrom <- min(gene.plotdata$start[gene.plotdata$gene_name %in% gene_to_plot]) - vis_win
    visto   <- max(gene.plotdata$end  [gene.plotdata$gene_name %in% gene_to_plot]) + vis_win
  }

  
  ggann <- plot.genome.region(visfrom = visfrom, visto = visto, 
                     genome=genome, plot.data=plot.data, gene.plotdata=gene.plotdata, TR.merged.data=TR.merged.data.gene,
             geom = geom_bar(), geom2 = NULL,
             prime='prime5', crop.FALSE=F, samples=NA, comb.all.samples=NA,
             sum.counts.in.window=T, bin_width = 200, sum.fun='sum', add.all.pos=F, y.thresh=0, minus.strand.down=T,
             #what.to.plot = c('coverage', 'gene annotation', 'transcripts'), 
             add.coverage=F, add.transcripts.plot=T, add.genome.plot=T, flip.panels=F, genome.only=F, transcripts.only=F,
             genome.and.transcripts=T, transcript.plot.scale=2,
             add.cageTSS=F, cagefr.clust=NULL, cagefr.clust.dist=6, asterisk.size = 2,
             
             gene_name_col = 'gene_name', 
             gene.label=T, add.unstranded=T, add.feature=T, 
             #gene.feature.width.thresh=700, flip.gene.y=T, 
             force.gene.y=F,
             #force.one.lab.per.gene=T, 
             force.all.gene.up.and.down=T,
             force.all.gene.down=F, angle=angle, 
             
             transcript.sizes=transcript.sizes, tr.size.multip = 0.75,
             sizes=sizes, gene.sizes=gene.sizes, gene.label.col='black', alpha=alpha, plot.title=NA, 
             palette=palette[1:3], tr.palette = colorvec, # palette[c(15,18)],
             tolower=F, vline=NULL, vline2=NULL, breakseq=breakseq, margins=NULL, #unit(c(1,1,1,1), "cm"),
             scales=scales, genomplot.scale=5.5, ylim=c(0, NA), y.log10=F, y.log2=F, ylims.gene=ylims.gene, y.multip=1.5, 
             ybreaks=NULL, ybreak_n=10,
             strip.text =  element_text(angle = 45, size = sizes*1.5, hjust = 0.5), facet_space=1,
             gene.aes = aes(xmin = start, xmax = end, y = ymin, fill = strand, forward = orientation, label = gene, fontface='bold'),
             facet_cropF = NULL, # facet_nested(rows = vars(hpi), cols=NULL, scales ),
             labels=NULL, legend.position.prime = 'top', theme_general = theme_bw(), 
             return.plot.data=F
             , ...
             )
             
  
  ggann

}


#ann.plotfun(visfrom = visfrom, visto = visto, gene_to_plot, orf.sum.perc, vis_win)

```


### Isoform Ratio plotting function
```{r}
### 

max_ncol   <- 6
row_multip <- 6
col_multip <- 4

plotfun <- function(DT, 
                    geom1 = geom_point(aes(x=Time, y=norm_read_count, colour = Ref_Class)),
                    #geom1 = geom_pointrange(aes(x=Time, y=mean, ymin = mean-sd, ymax = mean+sd, colour = Ref_Class)),
                    geom2 = geom_smooth(aes(x=Time, y=norm_read_count, color=Ref_Class, group = Ref_Class)),
                    geom3 = NULL,
                    ncol = 10, title = NULL, #"Normalized ORF counts", 
                    strip.backgr = pal_d3()(10)[2], palette=colorvec[],
                    save.plot=T, filename='plot.jpg', figw=12, figh=9,
                    legend.position= 'right', 
                    legend.name    = '',
                    TR_ratio=T,
                    ...) {
  
  if (nrow(DT) == 0) { 
    ggp <- NULL } else {   # try({ 
      
    gene_N <- length(unique(DT[,gene]))
  
    ggp <- 
      ggplot(DT) + 
      geom1 +
      geom2 + 
      geom3 +
      
      ## Color and Fill Scales
      #guides( color = guide_legend("Ref. Isoform") ) +
      scale_color_manual(values = palette, name=legend.name) +
      scale_fill_manual( values = palette, name=legend.name) +
     
      ## Axis limits and labels
      { if(TR_ratio) 
        scale_y_continuous(name = 'Transcript Isoform ratio', limits = c(0, 1))
      } +
      scale_x_continuous(name = 'Hours past infection') +
      
      #coord_cartesian(ylim = c(0, NULL)) +
      
      
      ## Theme elements
      theme_bw() +
      theme( text = element_text(size = 8),
             legend.position = legend.position, 
             plot.title = element_text(face = "bold"),
            # plot.margin = unit(c(1,1,1,1), 'mm'),
            # strip.background = element_rect(fill = alpha(strip.backgr, 0.4))
            #margins=unit(c(1,1,1,1), "cm"),
            ...) +
      
      ## Faceting
      #facet_wrap(~ gene, scales = 'free', ncol = ncol) +
      #facet_nested_wrap(~ gene, scales = 'free', ncol = ncol) +
      #facet_nested(cols=vars(gene, cell_line), scales = 'free', independent=T) + 
      
      ## Plot title
      ggtitle(title)
   
  if (gene_N < ncol) { 
    #ggp <- cowplot::plot_grid(ggp, NULL, rel_widths = c(gene_N, ncol - gene_N ), nrow=1)
  }
  
  }   # })
  
  if(save.plot) {
    message('saving plot to: ', filename, '...')
    ggsave(filename, ggp, width = figw, height = figh)
  }
  
  return(ggp)
 
}


strip.background <- palette[10]
```


### Merge IF ratio and annotation plots function
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}
#DT <- orf.sum.perc

mapfun <- function(gene_to_plot=NULL,  DT=orf.sum.perc, TRs_to_plot=NULL, recalc = F,
                   TR.annot.data = TR.merged.data.gene,
                   visfrom = NULL, visto = NULL, strand_to_plot = c('+', '-'), omit_other_genes=F,
                   rel_widths = c(5, 5), what2plot = 'annot_and_abund', save.plot=F, 
                   use_gene_norm_count=T, fscale='fixed', fnrow=1, 
                   breakseq=5000, min_TIF_ratio = NULL,
                   ...) { 
  
    stopifnot(!(is.null(gene_to_plot) & is.null(TRs_to_plot)))
  
  
    filename <- paste0(fig.dir, '/', 'IF.ratio_and_annotation_', gene_to_plot, '_', 'replicates_loess.jpg')
  
    if(!is.null(gene_to_plot) & is.null(TRs_to_plot)) {
      DTsub    <- DT[gene %in% gene_to_plot]
    } else if (is.null(gene_to_plot) & !is.null(TRs_to_plot)) {
      DTsub    <- DT[Ref_Class %in% TRs_to_plot]
    }
    
    DTsub    <- DTsub[strand %in% strand_to_plot]
    
    ## recalculate ratios without gene assignment, but with the filtered transcripts
    TR_ratio <- T
    
    if (recalc) {
      DTsub[, sum_read_count  := sum(read_count), by=.(sample, Time)]
      DTsub[, norm_read_count := read_count / sum_read_count]
      DTsub[is.na(norm_read_count), norm_read_count := 0]
      DTsub[, mean            := mean(norm_read_count, na.rm=T), by=.(Time, Ref_Class)]
      DTsub[, sd := sd(norm_read_count, na.rm=T), by=.(Time, Ref_Class)]
      
    ## Or use the the already normalized counts (or raw counts, but not neccessarly the isoform-ratios) 
    } else if (!use_gene_norm_count) {
      DTsub[, norm_base       := sum(read_count), by=sample]
      DTsub[, norm_read_count := read_count / norm_base]
      DTsub[, mean            := mean(norm_read_count, na.rm=T), by=.(Time, gene, Ref_Class)]
      DTsub[, sd              := sd(  norm_read_count, na.rm=T), by=.(Time, gene, Ref_Class)]
      TR_ratio <- F
      
    } else if (recalc & use_gene_norm_count & is.null(TRs_to_plot) ) {
    ## recalculate ratios with gene assignment AND with the filtered transcripts
      DTsub[, sum_read_count  := sum(read_count), by=.(sample, gene, Time)]
      DTsub[, norm_read_count := read_count / sum_read_count]
      DTsub[is.na(norm_read_count), norm_read_count := 0]
      DTsub[, mean            := mean(norm_read_count, na.rm=T), by=.(Time, gene, Ref_Class)]
      DTsub[, sd := sd(norm_read_count, na.rm=T), by=.(Time, gene, Ref_Class)]
      
    } else {
    ### OR use the gene-normalized transcript isoform ratios, calculated previously
    }
    
    ### filter by minimum transcript isoform ratio?
    if(!is.null(min_TIF_ratio)) {
      DTsub <- DTsub[ norm_read_count >= min_TIF_ratio, ]
    } 
    
    min_TIF_ratio_TRs <- unique(DTsub[,Ref_Class])
    
    ### Make the plot
    ggrt  <- plotfun(DTsub, 
                    geom1 = geom_linerange(aes(x=Time, y=mean, ymin = mean-sd, ymax = mean+sd)),
                    geom3 = geom_point(aes(x=Time, y=mean, fill = Ref_Class), shape=21, size=3),
                    geom2 = geom_line(aes(x=Time, y=mean, color=Ref_Class, group = Ref_Class), linewidth = 0.85),
                    #geom2 = geom_pointrange(aes(x=Time, y=mean, ymin = mean-sd, ymax = mean+sd, colour = Ref_Class)),
                    #geom3 = geom_smooth(aes(x=Time, y=mean, color=Ref_Class, group = Ref_Class), se = F, method = 'gam'),
                    save.plot = F,
                    TR_ratio  = TR_ratio,
                    ...) 
    
    if (is.null(gene_to_plot) & !is.null(TRs_to_plot)) { 
      ggrt <- ggrt + facet_wrap(~gene, nrow=fnrow, scales=fscale)
    }
  
    
    ggtr <- NULL
    try({
      
      
      
      if(!is.null(gene_to_plot) & is.null(TRs_to_plot)) {
        
        ## Filter for strands
        TR.annot.data.filt <- TR.annot.data[strand %in% strand_to_plot]
        
        ## Make other gene's transcripts NA
        TR.annot.data.filt[!gene %in% gene_to_plot, Ref_Class := NA]
        
        ## Completely omit other genes from the plot?
        if(omit_other_genes) {
          TR.annot.data.filt <- TR.annot.data.filt[gene %in% gene_to_plot, ]
        }
        
        if(is.null(visfrom) & is.null(visto)) {
          visfrom <- min(TR.annot.data.filt[gene %in% gene_to_plot, start]) - vis_win
          visto   <- max(TR.annot.data.filt[gene %in% gene_to_plot, end  ]) + vis_win
        }
      
      } else if (is.null(gene_to_plot) & !is.null(TRs_to_plot)) {
        TR.annot.data.filt <- TR.annot.data[Ref_Class %in% TRs_to_plot]
      }
      
      if(!is.null(min_TIF_ratio)) {
        TR.annot.data.filt <- TR.annot.data.filt[Ref_Class %in% min_TIF_ratio_TRs, ]
      } 
   
      ggtr <- ann.plotfun(gene_to_plot, vis_win=vis_win, 
                          visfrom = visfrom, visto = visto,
                          TR.merged.data = TR.annot.data.filt,
                          annot_fill_column = 'Ref_Class',
                          add.TR.labels = T, breakseq=breakseq)
    })
    
    
    if (what2plot == 'annot_and_abund') {
    
      if (is.null('ggtr')) {
        ggcomb <- cowplot::plot_grid(ggrt, ggtr, nrow = 1, rel_widths = c(3, 0.5), align = 'v', axis = 'tb')
      } else {
        ggcomb <- cowplot::plot_grid(ggrt, ggtr, nrow = 1, rel_widths = rel_widths, align = 'v', axis = 'tb')
      }
      
    } else if (what2plot == 'annot_only') {
      
      ggcomb <- ggtr
      
    } else if (what2plot == 'abund_only') {
      
      ggcomb <- ggrt
      
    }
      
    if(save.plot) { ggsave(filename, ggcomb, width = figw, height = figh) }
    
    ggcomb
  
  
}



#purrr::map(unique(orf.sum.perc[,gene]), mapfun, orf.sum.perc)
```


# START PLOTTING


## REVIEW


### Figure X - Anomalous transcripts

ORF23, ORF50, ORF67-ORF68, ORF64

```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

dir.create('Figures')

Fig.abunds1 <- cowplot::plot_grid(plotlist = list(
  mapfun('ORF23', orf.sum.perc, legend.position='right', title = 'A', what2plot = 'abund_only',
         recalc=F,
         breakseq = 1500,
         strand_to_plot = '+'
         #visfrom = min(TR.merged.data.gene[gene == 'ORF45' & strand == '+', start]) - vis_win,
         #visto   = max(TR.merged.data.gene[gene == 'ORF45' & strand == '+', end  ]) + vis_win
         ),
  mapfun('ORF50', orf.sum.perc, legend.position='right', title = 'B', what2plot = 'abund_only',
         recalc=F,
         breakseq = 1500,
         strand_to_plot = '+',
         visfrom = min(TR.merged.data.gene[gene == 'ORF50' & strand == '+', start]) - vis_win,
         visto   = max(TR.merged.data.gene[gene == 'ORF50' & strand == '+', end  ]) + vis_win
         )
), ncol = 1, align = 'v', axis = 'tlbr')


Fig.annots1 <- cowplot::plot_grid(plotlist = list(
  mapfun('ORF23', orf.sum.perc, what2plot = 'annot_only',
         recalc=F,
         breakseq = 1500,
         strand_to_plot = '+',
         #visfrom = min(TR.merged.data.gene[gene == 'ORF45' & strand == '+', start]) - vis_win,
         #visto   = max(TR.merged.data.gene[gene == 'ORF45' & strand == '+', end  ]) + vis_win
         ),
  mapfun('ORF50', orf.sum.perc, what2plot = 'annot_only',
         recalc=F,
         breakseq = 1500,
         strand_to_plot = '+',
         #visfrom = min(TR.merged.data.gene[gene == 'ORF50' & strand == '+', start]) - vis_win,
         #visto   = max(TR.merged.data.gene[gene == 'ORF50' & strand == '+', end  ]) + vis_win
         )
), ncol = 1, align = 'v', axis = 'tlbr')


Fig.isoforms1 <- cowplot::plot_grid(Fig.abunds1, Fig.annots1, ncol=2, rel_widths = c(1, 0.65), align = 'vh', axis = 'l')



Fig.abunds2 <- cowplot::plot_grid(plotlist = list(
  mapfun(c('ORF67', 'ORF68'), orf.sum.perc, legend.position='right', title = 'C', what2plot = 'abund_only',
         recalc=T,
         breakseq = 1500,
         strand_to_plot = '-',
         #visfrom = min(TR.merged.data.gene[gene == 'ORF67' & strand == '-', start]) - vis_win,
         #visto   = max(TR.merged.data.gene[gene == 'ORF67' & strand == '-', end  ]) + vis_win
         ),
  mapfun('ORF64', orf.sum.perc, legend.position='right', title = 'D', what2plot = 'abund_only',
         recalc=F,
         breakseq = 1500,
         #strand_to_plot = '+',
         visfrom = 111600, # min(TR.merged.data.gene[gene == 'ORF64' & strand == '-', start]) - vis_win,
         visto   = 118900  # max(TR.merged.data.gene[gene == 'ORF64' & strand == '-', end  ]) + vis_win
         )
), ncol = 1, align = 'vh', axis = 'tlbr')


Fig.annots2 <- cowplot::plot_grid(plotlist = list(
  mapfun(c('ORF67', 'ORF68'), orf.sum.perc, what2plot = 'annot_only',
         recalc=T,
         breakseq = 1500,
         strand_to_plot = '-',
         #visfrom = min(TR.merged.data.gene[gene == 'ORF67' & strand == '-', start]) - vis_win,
         #visto   = max(TR.merged.data.gene[gene == 'ORF67' & strand == '-', end  ]) + vis_win
         ),
  mapfun('ORF64', orf.sum.perc, what2plot = 'annot_only',
         recalc=F,
         breakseq = 1500,
         #strand_to_plot = '-',
         visfrom = 111600, # min(TR.merged.data.gene[gene == 'ORF64' & strand == '-', start]) - vis_win,
         visto   = 118900  # max(TR.merged.data.gene[gene == 'ORF64' & strand == '-', end  ]) + vis_win
         )
), ncol = 1, align = 'vh', axis = 'tlbr')


Fig.isoforms2 <- cowplot::plot_grid(Fig.abunds2, Fig.annots2, ncol=2, rel_widths = c(1, 0.85), align = 'vh', axis = 'l')



Fig.isoforms <- cowplot::plot_grid(Fig.isoforms1, Fig.isoforms2, ncol=2, rel_widths = c(1, 0.85))

#ggsave('./Figures/Figure 7_REVIEW_ori.tif', Fig.isoforms, height = 12, width = 24, dpi = 300)
ggsave('./Figures/Figure X_Anomaluous transcripts.jpg', Fig.isoforms, height = 10, width = 20)

#Fig.isoforms
```



### Figure 7 - DTU transcripts -->> OK

ORF11, ORF13, ORF14, ORF19, ORF40, ORF54

```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

dir.create('Figures')

Fig.abunds1 <- cowplot::plot_grid(plotlist = list(
  mapfun('ORF11', orf.sum.perc, legend.position='right', title = 'A', what2plot = 'abund_only',
         recalc=F,
         strand_to_plot = '+',
         visfrom = min(TR.merged.data.gene[gene == 'ORF11' & strand == '+', start]) - vis_win,
         visto   = max(TR.merged.data.gene[gene == 'ORF11' & strand == '+', end  ]) + vis_win), 
  mapfun('ORF13', orf.sum.perc, legend.position='right', title = 'C', what2plot = 'abund_only',
         recalc=F,
         strand_to_plot = '+',
         breakseq = 1500,
         visfrom  = 14980,    # min(TR.merged.data.gene[gene == 'ORF13' & strand == '+', start]) - vis_win,
         visto    = 21600 ),  # max(TR.merged.data.gene[gene == 'ORF13' & strand == '+', end  ]) + vis_win),
  mapfun('ORF14', orf.sum.perc, legend.position='right', title = 'E', what2plot = 'abund_only',
         recalc=F,
         strand_to_plot = '+',
         breakseq = 1500,
         visfrom  = 14980,    # min(TR.merged.data.gene[gene == 'ORF13' & strand == '+', start]) - vis_win,
         visto    = 21600 )   # max(TR.merged.data.gene[gene == 'ORF13' & strand == '+', end  ]) + vis_win),
  #mapfun('ORF19', orf.sum.perc, legend.position='right', title = 'D', what2plot = 'abund_only'),
  #mapfun('ORF40', orf.sum.perc, legend.position='right', title = 'E', what2plot = 'abund_only'),
  #mapfun('ORF54', orf.sum.perc, legend.position='right', title = 'F', what2plot = 'abund_only')
), ncol = 1, align = 'v', axis = 'tlbr', rel_heights = c(0.7, 0.8, 1))


Fig.annots1 <- cowplot::plot_grid(plotlist = list(
  mapfun('ORF11', orf.sum.perc, what2plot = 'annot_only',
         recalc=F,
         strand_to_plot = '+',
         visfrom = min(TR.merged.data.gene[gene == 'ORF11' & strand == '+', start]) - vis_win,
         visto   = max(TR.merged.data.gene[gene == 'ORF11' & strand == '+', end  ]) + vis_win),  
  mapfun('ORF13', orf.sum.perc, what2plot = 'annot_only',
         recalc=F,
         strand_to_plot = '+',
         breakseq = 1500,
         visfrom  = 14980,    # min(TR.merged.data.gene[gene == 'ORF13' & strand == '+', start]) - vis_win,
         visto    = 21600 ),  # max(TR.merged.data.gene[gene == 'ORF13' & strand == '+', end  ]) + vis_win),
  mapfun('ORF14', orf.sum.perc, what2plot = 'annot_only',
         recalc=F,
         strand_to_plot = '+',
         breakseq = 1500,
         visfrom  = 14980,    # min(TR.merged.data.gene[gene == 'ORF13' & strand == '+', start]) - vis_win,
         visto    = 21600 )  # max(TR.merged.data.gene[gene == 'ORF13' & strand == '+', end  ]) + vis_win),
  #mapfun('ORF19', orf.sum.perc, what2plot = 'annot_only'),
  #mapfun('ORF40', orf.sum.perc, what2plot = 'annot_only'),
  #mapfun('ORF54', orf.sum.perc, what2plot = 'annot_only')
), ncol = 1, align = 'v', axis = 'tlbr', rel_heights = c(0.7, 0.8, 1))


Fig.isoforms1 <- cowplot::plot_grid(Fig.abunds1, Fig.annots1, ncol=2, rel_widths = c(1, 0.65), align = 'vh', axis = 'l')



Fig.abunds2 <- cowplot::plot_grid(plotlist = list(
  #mapfun('ORF11', orf.sum.perc, legend.position='right', title = 'A', what2plot = 'abund_only'), 
  #mapfun('ORF13', orf.sum.perc, legend.position='right', title = 'B', what2plot = 'abund_only'),
  #mapfun('ORF14', orf.sum.perc, legend.position='right', title = 'C', what2plot = 'abund_only'),
  mapfun('ORF19', orf.sum.perc, legend.position='right', title = 'B', what2plot = 'abund_only',
         recalc=F,
         strand_to_plot = '+',
         visfrom = min(TR.merged.data.gene[gene == 'ORF19' & strand == '+', start]) - vis_win,
         visto   = max(TR.merged.data.gene[gene == 'ORF19' & strand == '+', end  ]) + vis_win),
  mapfun('ORF40', orf.sum.perc, legend.position='right', title = 'D', what2plot = 'abund_only',
         recalc=F,
         strand_to_plot = '-',
         visfrom = min(TR.merged.data.gene[gene == 'ORF40' & strand == '-', start]) - vis_win,
         visto   = max(TR.merged.data.gene[gene == 'ORF40' & strand == '-', end  ]) + vis_win),
  mapfun('ORF54', orf.sum.perc, legend.position='right', title = 'F', what2plot = 'abund_only',
         recalc=F,
         strand_to_plot = '+',
         visfrom = min(TR.merged.data.gene[gene == 'ORF54' & strand == '+', start]) - vis_win,
         visto   = max(TR.merged.data.gene[gene == 'ORF54' & strand == '+', end  ]) + vis_win)
), ncol = 1, align = 'vh', axis = 'tlbr', rel_heights = c(0.7, 0.8, 1))


Fig.annots2 <- cowplot::plot_grid(plotlist = list(
  #mapfun('ORF11', orf.sum.perc, what2plot = 'annot_only'),  
  #mapfun('ORF13', orf.sum.perc, what2plot = 'annot_only'),
  #mapfun('ORF14', orf.sum.perc, what2plot = 'annot_only'),
  mapfun('ORF19', orf.sum.perc, what2plot = 'annot_only',
         recalc=F,
         strand_to_plot = '+',
         visfrom = min(TR.merged.data.gene[gene == 'ORF19' & strand == '+', start]) - vis_win,
         visto   = max(TR.merged.data.gene[gene == 'ORF19' & strand == '+', end  ]) + vis_win),
  mapfun('ORF40', orf.sum.perc, what2plot = 'annot_only',
         recalc=F,
         strand_to_plot = '-',
         visfrom = min(TR.merged.data.gene[gene == 'ORF40' & strand == '-', start]) - vis_win,
         visto   = max(TR.merged.data.gene[gene == 'ORF40' & strand == '-', end  ]) + vis_win),
  mapfun('ORF54', orf.sum.perc, what2plot = 'annot_only',
         recalc=F,
         strand_to_plot = '+',
         visfrom = min(TR.merged.data.gene[gene == 'ORF54' & strand == '+', start]) - vis_win,
         visto   = max(TR.merged.data.gene[gene == 'ORF54' & strand == '+', end  ]) + vis_win)
), ncol = 1, align = 'vh', axis = 'tlbr', rel_heights = c(0.7, 0.8, 1))


Fig.isoforms2 <- cowplot::plot_grid(Fig.abunds2, Fig.annots2, ncol=2, rel_widths = c(1, 0.85), align = 'vh', axis = 'l')



Fig.isoforms <- cowplot::plot_grid(Fig.isoforms1, Fig.isoforms2, ncol=2, rel_widths = c(1, 0.85))

#ggsave('./Figures/Figure 7_REVIEW_ori.tif', Fig.isoforms, height = 12, width = 24, dpi = 300)
ggsave('./Figures/Figure DTU REVIEW.jpg', Fig.isoforms, height = 14, width = 24)

#Fig.isoforms
```



### Figure 3 - Spliced Transcripts -->> OK

ORF8, ORF9, ORF53, ORF38, ORF58, NOIR 

```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}


Fig.abunds1 <- cowplot::plot_grid(plotlist = list(
  mapfun('ORF38', orf.sum.perc, legend.position='right', title = 'A', what2plot = 'abund_only',
         breakseq=1500,
         recalc=T,
         strand_to_plot = '+',
         visfrom = 65085,
         visto   = 71050),
  mapfun('ORF9', orf.sum.perc, legend.position='right', title = 'C', what2plot = 'abund_only',
         breakseq=2000,
         recalc=T, #min_TIF_ratio = 0.01,
         strand_to_plot = '-',
         visfrom = 10400, #min(TR.merged.data.gene[gene == 'ORF9' & strand == '-', start]) - vis_win,
         visto   = 16600 #max(TR.merged.data.gene[gene == 'ORF9' & strand == '-', end  ]) + vis_win
         ),
  mapfun('ORF58', orf.sum.perc, legend.position='right', title = 'E', what2plot = 'abund_only',
         breakseq=1500,
         recalc=T,
         strand_to_plot = '+',
         visfrom = 96600,
         visto   = 106000
         )
), ncol = 1, align = 'vh', axis = 'tlbr', rel_heights = c(0.8, 0.8, 1))



Fig.annots1 <- cowplot::plot_grid(plotlist = list(
  mapfun('ORF38', orf.sum.perc, what2plot = 'annot_only',
         breakseq=1500,
         recalc=T,
         strand_to_plot = '+',
         visfrom = 65085,
         visto   = 71050),
  mapfun('ORF9', orf.sum.perc, what2plot = 'annot_only',
         breakseq=2000,
         recalc=T,  #min_TIF_ratio = 0.01,
         strand_to_plot = '-',
         visfrom = 10400, #min(TR.merged.data.gene[gene == 'ORF9' & strand == '-', start]) - vis_win,
         visto   = 16600 #max(TR.merged.data.gene[gene == 'ORF9' & strand == '-', end  ]) + vis_win
         ),
  mapfun('ORF58', orf.sum.perc, what2plot = 'annot_only',
         breakseq=1500,
         recalc=T,
         strand_to_plot = '+',
         visfrom = 96600,
         visto   = 106000
         )
), ncol = 1, align = 'vh', axis = 'tlbr', rel_heights = c(0.8, 0.8, 1))


Fig.isoforms1 <- cowplot::plot_grid(Fig.abunds1, Fig.annots1, ncol=2, rel_widths = c(1.8, 1))



Fig.abunds2 <- cowplot::plot_grid(plotlist = list(
  mapfun('ORF8', orf.sum.perc, legend.position='right', title = 'B', what2plot = 'abund_only',
         breakseq=1000,
         recalc=T,
         strand_to_plot = '+',
         visfrom = 8900, #min(TR.merged.data.gene[gene == 'ORF9' & strand == '-', start]) - vis_win,
         visto   = 11100 #max(TR.merged.data.gene[gene == 'ORF9' & strand == '-', end  ]) + vis_win
         ),
  mapfun('NOIR', orf.sum.perc, legend.position='right', title = 'D', what2plot = 'abund_only',
         breakseq=1000,
         recalc=T,
         strand_to_plot = '+',
         visfrom = 118400, # min(TR.merged.data.gene[gene == 'NOIR' & strand == '+', start]) - vis_win,
         visto   = 120100  # max(TR.merged.data.gene[gene == 'NOIR' & strand == '+', end  ]) + vis_win
         ),
  mapfun(c('ORF53'), orf.sum.perc, legend.position='right', title = 'F', what2plot = 'abund_only',
         breakseq=1500,
         recalc=F,
         #strand_to_plot = '+',
         visfrom = 92600,
         visto   = 99600)
), ncol = 1, align = 'vh', axis = 'tlbr', rel_heights = c(0.8, 0.8, 1))



Fig.annots2 <- cowplot::plot_grid(plotlist = list(
  mapfun('ORF8', orf.sum.perc, what2plot = 'annot_only',
         breakseq=1000,
         recalc=T,
         strand_to_plot = '+',
         visfrom = 8900, #min(TR.merged.data.gene[gene == 'ORF9' & strand == '-', start]) - vis_win,
         visto   = 11100 #max(TR.merged.data.gene[gene == 'ORF9' & strand == '-', end  ]) + vis_win
         ),
  mapfun('NOIR', orf.sum.perc, what2plot = 'annot_only',
         breakseq=1000,
         recalc=T,
         strand_to_plot = '+',
         visfrom = 118400, # min(TR.merged.data.gene[gene == 'NOIR' & strand == '+', start]) - vis_win,
         visto   = 120100  # max(TR.merged.data.gene[gene == 'NOIR' & strand == '+', end  ]) + vis_win
         ),
  mapfun(c('ORF53'), orf.sum.perc, what2plot = 'annot_only',
         breakseq=1500,
         recalc=F,
         #strand_to_plot = '+',
         visfrom = 92600,
         visto   = 99600)
), ncol = 1, align = 'vh', axis = 'tlbr', rel_heights = c(0.8, 0.8, 1))


Fig.isoforms2 <- cowplot::plot_grid(Fig.abunds2, Fig.annots2, ncol=2, rel_widths = c(1.4, 1))



Fig.splice <- cowplot::plot_grid(Fig.isoforms1, Fig.isoforms2, ncol=2, rel_widths = c(1.15, 1))


#ggsave('./Figures/Figure 3_REVIEW_ori.tif', Fig.splice, height = 10, width = 24, dpi = 300)
ggsave('./Figures/Figure Spliced Transcripts REVIEW.jpg', Fig.splice, height = 14, width = 24)

#Fig.splice
```



###  Supp Figure S3 - Spliced Transcripts 2 -->> OK

ORF47-44, ORF65
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}
Fig.abunds1 <- cowplot::plot_grid(plotlist = list(
  mapfun('ORF65',  orf.sum.perc, legend.position='right', title = 'A', what2plot = 'abund_only',
         breakseq=1000,
         recalc=T,
         strand_to_plot = '-'
         #visfrom = min(TR.merged.data.gene[gene == 'ORF65' & strand == '-', start]) - vis_win,
         #visto   = max(TR.merged.data.gene[gene == 'ORF65' & strand == '-', end  ]) + vis_win
         ),
  mapfun(c('ORF44'), orf.sum.perc, legend.position='right', title = 'B', what2plot = 'abund_only',
         breakseq=1500,
         recalc=T, #min_TIF_ratio = 0.01
         #strand_to_plot = '-',
         visfrom = 83000, # min(TR.merged.data.gene[gene == 'ORF9' & strand == '-', start]) - vis_win,
         visto   = 93000  # max(TR.merged.data.gene[gene == 'ORF9' & strand == '-', end  ]) + vis_win
         )
), ncol = 1, align = 'vh', axis = 'tlbr', rel_heights = c(0.8, 1))


Fig.annots1 <- cowplot::plot_grid(plotlist = list(
  mapfun('ORF65', orf.sum.perc, what2plot = 'annot_only',
         breakseq=1000,
         recalc=T,
         strand_to_plot = '-'
         #visfrom = 65085,
         #visto   = 71050
         ),
  mapfun(c('ORF44'), orf.sum.perc, what2plot = 'annot_only',
         breakseq=1500,
         recalc=T,  #min_TIF_ratio = 0.01
         #strand_to_plot = '-',
         visfrom = 83000, # min(TR.merged.data.gene[gene == 'ORF9' & strand == '-', start]) - vis_win,
         visto   = 93000  # max(TR.merged.data.gene[gene == 'ORF9' & strand == '-', end  ]) + vis_win
         )
), ncol = 1, align = 'vh', axis = 'tlbr', rel_heights = c(0.8, 1))


Fig.isoforms1 <- cowplot::plot_grid(Fig.abunds1, Fig.annots1, ncol=2, rel_widths = c(2, 1))





Fig.splice <- cowplot::plot_grid(Fig.isoforms1#, Fig.isoforms2, ncol=2, rel_widths = c(1, 1.4)
                                 )


#ggsave('./Figures/Figure 3_REVIEW_ori.tif', Fig.splice, height = 10, width = 24, dpi = 300)
ggsave('./Figures/Supp Fig Spliced 2 REVIEW.jpg', Fig.splice, height = 12, width = 16)

#Fig.splice
```





## For each gene

```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=T}

dir.create('Figures/GeneIsoforms')

genes_to_plot <- unique(orf.sum.perc$gene)

for(gene_to_plot in genes_to_plot) {

  message('plotting isoforms of gene: ', gene_to_plot)
  try({
    Fig.isoform <- cowplot::plot_grid(
      
        mapfun(gene_to_plot,  orf.sum.perc, what2plot = 'abund_only',
               breakseq=1000,
               #recalc=T,
               #strand_to_plot = '-'
               #visfrom = min(TR.merged.data.gene[gene == 'ORF65' & strand == '-', start]) - vis_win,
               #visto   = max(TR.merged.data.gene[gene == 'ORF65' & strand == '-', end  ]) + vis_win
               ),
        mapfun(gene_to_plot, orf.sum.perc, what2plot = 'annot_only',
               breakseq=1500,
               #recalc=T,  #min_TIF_ratio = 0.01
               #strand_to_plot = '-',
               #visfrom = 83000, # min(TR.merged.data.gene[gene == 'ORF9' & strand == '-', start]) - vis_win,
               #visto   = 93000  # max(TR.merged.data.gene[gene == 'ORF9' & strand == '-', end  ]) + vis_win
               ), 
        nrow = 1, align = 'vh', axis = 'tlbr', rel_heights = c(0.8, 1))
    
  
    ggsave(paste0('Figures/GeneIsoforms/', gene_to_plot, '_isoforms.jpg'), Fig.isoform, 
           height = 6, width = 12)
  })
}



```







## Tests

### V1
```{r, fig.show='hold',  fig.width = 12, fig.height = 6, eval=F}

gene_to_plot <- 'ORF38'


Fig.test <- cowplot::plot_grid(
  mapfun(gene_to_plot=gene_to_plot,  orf.sum.perc, 
       TR.annot.data = TR.merged.data.gene,
       strand_to_plot = '+',
       #visfrom = min(TR.merged.data.gene[gene == gene_to_plot & strand == '+', start]) - vis_win,
       #visto   = max(TR.merged.data.gene[gene == gene_to_plot & strand == '+', end  ]) + vis_win,
       visfrom =NULL, visto =NULL,
       legend.position='right',  rel_widths=c(1, 1), title = gene_to_plot, what2plot = 'abund_only'),
 mapfun(gene_to_plot=gene_to_plot,  orf.sum.perc, 
       TR.annot.data = TR.merged.data.gene,
       strand_to_plot = '+',
       #visfrom = min(TR.merged.data.gene[gene == gene_to_plot & strand == '+', start]) - vis_win,
       #visto   = max(TR.merged.data.gene[gene == gene_to_plot & strand == '+', end  ]) + vis_win,
       visfrom =NULL, visto =NULL,
       legend.position='right',  rel_widths=c(1, 1), title = gene_to_plot, what2plot = 'annot_only'),
  ncol=2, rel_widths = c(1, 0.65), align = 'vh', axis = 'l')


###
gene_to_plot <- 'ORF38'
visfrom <- min(TR.merged.data.gene[gene == gene_to_plot & strand == '+', start])
visto   <- max(TR.merged.data.gene[gene == gene_to_plot & strand == '+', end  ])


TRs_to_plot <- TR.merged.data.gene[,.(seqnames, start, end, Ref_Class)]
TRs_to_plot <- genome_join(TRs_to_plot, data.frame(seqnames=genome, start=visfrom, end=visto))
TRs_to_plot <- unique(TRs_to_plot$Ref_Class)

Fig.test <- cowplot::plot_grid(
  mapfun(gene_to_plot=NULL,  orf.sum.perc, TRs_to_plot=TRs_to_plot, 
       TR.annot.data = TR.merged.data.gene,
       strand_to_plot = '+',
       recalc=F,  use_gene_norm_count=F, fscale='free', 
       #visfrom = min(TR.merged.data.gene[gene == gene_to_plot & strand == '+', start]) - vis_win,
       #visto   = max(TR.merged.data.gene[gene == gene_to_plot & strand == '+', end  ]) + vis_win,
       visfrom =visfrom, visto =visto,
       legend.position='bottom',  rel_widths=c(1, 1), title = gene_to_plot, what2plot = 'abund_only'),
 mapfun(gene_to_plot=NULL,  orf.sum.perc, TRs_to_plot=TRs_to_plot,
       TR.annot.data = TR.merged.data.gene,
       strand_to_plot = '+',
       recalc=T,
       #visfrom = min(TR.merged.data.gene[gene == gene_to_plot & strand == '+', start]) - vis_win,
       #visto   = max(TR.merged.data.gene[gene == gene_to_plot & strand == '+', end  ]) + vis_win,
       visfrom =visfrom, visto =visto, 
       legend.position='bottom',  rel_widths=c(1, 1), title = gene_to_plot, what2plot = 'annot_only'),
  ncol=2, rel_widths = c(2, 0.85), align = 'vh', axis = 'l')




ggsave('Figures/ORF38_TRs.jpg', Fig.test, width = 12, height = 7)
Fig.test


###
gene_to_plot <- 'ORF50'
visfrom <- min(TR.merged.data.gene[gene == gene_to_plot & strand == '+', start])
visto   <- max(TR.merged.data.gene[gene == gene_to_plot & strand == '+', end  ])


TRs_to_plot <- TR.merged.data.gene[,.(seqnames, start, end, Ref_Class)]
TRs_to_plot <- genome_join(TRs_to_plot, data.frame(seqnames=genome, start=visfrom, end=visto))
TRs_to_plot <- unique(TRs_to_plot$Ref_Class)

vis_win <- 1500
visfrom <- min(TR.merged.data.gene[gene == gene_to_plot & strand == '+', start]) - vis_win*2
visto   <- max(TR.merged.data.gene[gene == gene_to_plot & strand == '+', end  ]) + vis_win


Fig.test <- cowplot::plot_grid(
  mapfun(gene_to_plot=NULL,  orf.sum.perc, TRs_to_plot=TRs_to_plot, 
       TR.annot.data = TR.merged.data.gene,
       strand_to_plot = '+',
       recalc=F,  use_gene_norm_count=F, fscale='free', fnrow = 2,
       #visfrom = min(TR.merged.data.gene[gene == gene_to_plot & strand == '+', start]) - vis_win,
       #visto   = max(TR.merged.data.gene[gene == gene_to_plot & strand == '+', end  ]) + vis_win,
       visfrom =visfrom, visto =visto,
       legend.position='bottom',  rel_widths=c(1, 1), title = gene_to_plot, what2plot = 'abund_only'),
 mapfun(gene_to_plot=NULL,  orf.sum.perc, TRs_to_plot=TRs_to_plot,
       TR.annot.data = TR.merged.data.gene,
       strand_to_plot = '+',
       recalc=T,
       #visfrom = min(TR.merged.data.gene[gene == gene_to_plot & strand == '+', start]) - vis_win,
       #visto   = max(TR.merged.data.gene[gene == gene_to_plot & strand == '+', end  ]) + vis_win,
       visfrom =visfrom, visto =visto, 
       legend.position='bottom',  rel_widths=c(1, 1), title = gene_to_plot, what2plot = 'annot_only'),
  ncol=2, rel_widths = c(2, 0.85), align = 'vh', axis = 'l')




ggsave('Figures/ORF50_TRs.jpg', Fig.test, width = 18, height = 10)
Fig.test


```
### V2
```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=F}

ORFs_toplot <- c("ORF38", "ORF45", "ORF50", "ORF54", "ORF67", "ORF70", "ORF75")


###
gene_to_plot <- 'ORF67'
for (gene_to_plot in ORFs_toplot) {
  
  visfrom <- min(TR.merged.data.gene[gene == gene_to_plot & strand == '+', start])
  visto   <- max(TR.merged.data.gene[gene == gene_to_plot & strand == '+', end  ])
  
  ## Plot and recalculate abundnance of each transcript in the region
  TRs_to_plot <- TR.merged.data.gene[,.(seqnames, start, end, Ref_Class)]
  TRs_to_plot <- genome_join(TRs_to_plot, data.frame(seqnames=genome, start=visfrom, end=visto))
  TRs_to_plot <- unique(TRs_to_plot$Ref_Class)
  
  ## Plot and recalculate abundnance of each transcript from the GENE ONLY
  #TRs_to_plot <- unique(TR.merged.data.gene[gene == gene_to_plot, Ref_Class])
  
  vis_win <- 1500
  visfrom <- min(TR.merged.data.gene[gene == gene_to_plot & strand == '+', start]) - vis_win*2
  visto   <- max(TR.merged.data.gene[gene == gene_to_plot & strand == '+', end  ]) + vis_win
  
  
  Fig.TXs <- cowplot::plot_grid(
    mapfun(gene_to_plot=NULL,  orf.sum.perc, TRs_to_plot=TRs_to_plot, 
         TR.annot.data = TR.merged.data.gene,
         strand_to_plot = '+',
         recalc=F,  use_gene_norm_count=F, fscale='free', fnrow = 2,
         #visfrom = min(TR.merged.data.gene[gene == gene_to_plot & strand == '+', start]) - vis_win,
         #visto   = max(TR.merged.data.gene[gene == gene_to_plot & strand == '+', end  ]) + vis_win,
         visfrom =visfrom, visto =visto,
         legend.position='bottom',  rel_widths=c(1, 1), title = gene_to_plot, what2plot = 'abund_only'),
   mapfun(gene_to_plot=NULL,  orf.sum.perc, TRs_to_plot=TRs_to_plot,
         TR.annot.data = TR.merged.data.gene,
         strand_to_plot = '+',
         recalc=T,
         #visfrom = min(TR.merged.data.gene[gene == gene_to_plot & strand == '+', start]) - vis_win,
         #visto   = max(TR.merged.data.gene[gene == gene_to_plot & strand == '+', end  ]) + vis_win,
         visfrom =visfrom, visto =visto, 
         legend.position='bottom',  rel_widths=c(1, 1), title = gene_to_plot, what2plot = 'annot_only'),
    ncol=2, rel_widths = c(2, 0.85), align = 'vh', axis = 'l')
  

  ggsave(paste0('Figures/', gene_to_plot, '_TRs.jpg'), Fig.TXs, width = 18, height = 10)

  Fig.TXs
  
  strand_to_plot <- unique(TR.merged.data.gene[gene == gene_to_plot, strand])
  vis_win <- 1500
  visfrom <- min(TR.merged.data.gene[gene == gene_to_plot & strand %in% strand_to_plot, start]) - vis_win*2
  visto   <- max(TR.merged.data.gene[gene == gene_to_plot & strand %in% strand_to_plot, end  ]) + vis_win
  
  Fig.Gene <- cowplot::plot_grid(
    mapfun(gene_to_plot=gene_to_plot,  orf.sum.perc, TRs_to_plot=NULL, 
         TR.annot.data = TR.merged.data.gene,
         strand_to_plot = strand_to_plot,
         recalc=F,  use_gene_norm_count=F, fscale='free', fnrow = 2,
         #visfrom = min(TR.merged.data.gene[gene == gene_to_plot & strand == '+', start]) - vis_win,
         #visto   = max(TR.merged.data.gene[gene == gene_to_plot & strand == '+', end  ]) + vis_win,
         visfrom =visfrom, visto =visto,
         legend.position='bottom',  rel_widths=c(1, 1), title = gene_to_plot, what2plot = 'abund_only'),
   mapfun(gene_to_plot=gene_to_plot,  orf.sum.perc, TRs_to_plot=NULL,
         TR.annot.data = TR.merged.data.gene,
         strand_to_plot = strand_to_plot,
         recalc=T,
         #visfrom = min(TR.merged.data.gene[gene == gene_to_plot & strand == '+', start]) - vis_win,
         #visto   = max(TR.merged.data.gene[gene == gene_to_plot & strand == '+', end  ]) + vis_win,
         visfrom =visfrom, visto =visto, 
         legend.position='bottom',  rel_widths=c(1, 1), title = gene_to_plot, what2plot = 'annot_only'),
    ncol=2, rel_widths = c(2, 0.85), align = 'vh', axis = 'l')
  

  ggsave(paste0('Figures/', gene_to_plot, '_Gene.jpg'), Fig.Gene, width = 18, height = 10)

  Fig.Gene

  
}

```



## Archive

### Figure 3 V1 - Spliced Transcripts

ORF9, ORF38, ORF65, NOIR 

```{r, fig.show='hold',  fig.width = figw, fig.height = figh, eval=F}


Fig.abunds1 <- cowplot::plot_grid(plotlist = list(
  mapfun('NOIR', orf.sum.perc, legend.position='right', title = 'A', what2plot = 'abund_only',
         breakseq=1500,
         recalc=T,
         strand_to_plot = '+',
         visfrom = 118400, # min(TR.merged.data.gene[gene == 'NOIR' & strand == '+', start]) - vis_win,
         visto   = 120100  # max(TR.merged.data.gene[gene == 'NOIR' & strand == '+', end  ]) + vis_win
         ),
  mapfun('ORF9', orf.sum.perc, legend.position='right', title = 'B', what2plot = 'abund_only',
         breakseq=1500,
         recalc=T,
         strand_to_plot = '-',
         visfrom = 10400, #min(TR.merged.data.gene[gene == 'ORF9' & strand == '-', start]) - vis_win,
         visto   = 16600 #max(TR.merged.data.gene[gene == 'ORF9' & strand == '-', end  ]) + vis_win
         )
  #mapfun('ORF38', orf.sum.perc, legend.position='right', title = 'C', what2plot = 'abund_only')
  #mapfun('ORF65', orf.sum.perc, legend.position='right', title = 'D', what2plot = 'abund_only')
), ncol = 1, align = 'vh', axis = 'tlbr')


Fig.annots1 <- cowplot::plot_grid(plotlist = list(
  mapfun('NOIR', orf.sum.perc, what2plot = 'annot_only',
         breakseq=1500,
         recalc=T,
         strand_to_plot = '+',
         visfrom = 118400, # min(TR.merged.data.gene[gene == 'NOIR' & strand == '+', start]) - vis_win,
         visto   = 120100  # max(TR.merged.data.gene[gene == 'NOIR' & strand == '+', end  ]) + vis_win
         ),  
  mapfun('ORF9', orf.sum.perc, what2plot = 'annot_only',
         breakseq=1500,
         recalc=T,
         strand_to_plot = '-',
         visfrom = 10400, #min(TR.merged.data.gene[gene == 'ORF9' & strand == '-', start]) - vis_win,
         visto   = 16600 #max(TR.merged.data.gene[gene == 'ORF9' & strand == '-', end  ]) + vis_win
         )
  #mapfun('ORF38', orf.sum.perc, what2plot = 'annot_only'),
  #mapfun('ORF65', orf.sum.perc, what2plot = 'annot_only')
), ncol = 1, align = 'vh', axis = 'tlbr')


Fig.isoforms1 <- cowplot::plot_grid(Fig.abunds1, Fig.annots1, ncol=2, rel_widths = c(2, 1))




Fig.abunds2 <- cowplot::plot_grid(plotlist = list(
  #mapfun('NOIR', orf.sum.perc, legend.position='right', title = 'A', what2plot = 'abund_only'), 
  #mapfun('ORF9', orf.sum.perc, legend.position='right', title = 'B', what2plot = 'abund_only'),
  mapfun('ORF38', orf.sum.perc, legend.position='right', title = 'C', what2plot = 'abund_only',
         breakseq=1500,
         recalc=T,
         strand_to_plot = '+',
         visfrom = 65085,
         visto   = 71050),
  mapfun('ORF65',  orf.sum.perc, legend.position='right', title = 'D', what2plot = 'abund_only',
         breakseq=1500,
         recalc=T,
         strand_to_plot = '-',
         visfrom = min(TR.merged.data.gene[gene == 'ORF65' & strand == '-', start]) - vis_win,
         visto   = max(TR.merged.data.gene[gene == 'ORF65' & strand == '-', end  ]) + vis_win)
), ncol = 1, align = 'vh', axis = 'tlbr')


Fig.annots2 <- cowplot::plot_grid(plotlist = list(
  #mapfun('NOIR', orf.sum.perc, what2plot = 'annot_only'),  
  #mapfun('ORF9', orf.sum.perc, what2plot = 'annot_only'),
  mapfun('ORF38', orf.sum.perc, what2plot = 'annot_only',
         breakseq=1500,
         recalc=T,
         strand_to_plot = '+',
         visfrom = 65085,
         visto   = 71050),
  mapfun('ORF65',  orf.sum.perc, what2plot = 'annot_only',
         breakseq=1500,
         recalc=T,
         #strand_to_plot = '-',
         visfrom = min(TR.merged.data.gene[gene == 'ORF65' & strand == '-', start]) - vis_win,
         visto   = max(TR.merged.data.gene[gene == 'ORF65' & strand == '-', end  ]) + vis_win)
), ncol = 1, align = 'vh', axis = 'tlbr')


Fig.isoforms2 <- cowplot::plot_grid(Fig.abunds2, Fig.annots2, ncol=2, rel_widths = c(1.4, 1))



Fig.splice <- cowplot::plot_grid(Fig.isoforms1, Fig.isoforms2, ncol=2, rel_widths = c(1, 1.4))


#ggsave('./Figures/Figure 3_REVIEW_ori.tif', Fig.splice, height = 10, width = 24, dpi = 300)
ggsave('./Figures/Figure 3_REVIEW.jpg', Fig.splice, height = 10, width = 24)

#Fig.splice
```


```{r}
## Export data
if (writetables) {
  
  write.table(supp_tab_S2, 'Figures/Supp Table S2 - Transcript Annotaions and Read Counts REVIEW.tsv', 
              quote = F, row.names = F, sep = '\t')
  
  write.table(mean_abundance, paste0(fig.dir, '/', EndType, '_mean_abundance.tsv'), quote = F, row.names = F, sep = '\t')
  
  write.table(sd_abundance, paste0(fig.dir, '/', EndType, '_sd_abundance.tsv'), quote = F, row.names = F, sep = '\t')
  
  write.table(orf.sum.perc, paste0(fig.dir, '/', EndType, '_counts.norm.tsv'), quote = F, row.names = F, sep = '\t')
  
  write.table(orf.sum.gene.TX, paste0(fig.dir, '/', EndType, '_gene_counts.tsv'), quote = F, row.names = F, sep = '\t')

}
```



```{r bib, include=FALSE}
# KEEP THIS AT THE END OF THE DOCUMENT TO GENERATE A LOCAL bib FILE FOR PKGS USED
knitr::write_bib(sub("^package:", "", grep("package", search(), value=TRUE)), file='skeleton.bib')
```
